---
title: "lineages_pull"
author: "DSSU/DIQA"
date: "11/16/2021"
output: html_document
---


# Load R libraries
```{r}
library(reticulate)
library(tidyverse)
library(fs)
library(lubridate)
library(here)
library(sendmailR)
```

# Load python libraries
```{python}
import requests
```

# API Connection
```{python}
# API request
lineages_request = requests.get('https://raw.githubusercontent.com/cov-lineages/pango-designation/master/lineage_notes.txt')

# Grab content of webpage
lineages_webcontent = lineages_request.content

# Convert to string
lineages_webcontent_string = str(lineages_webcontent,'utf-8')
```

```{r}
# call lineages_webcontent_string from python environment to R environment
lineages_webcontent_string <- py$lineages_webcontent_string

# split by "\n" to get a list by row, unlist
lineages_string_by_row <- lineages_webcontent_string %>%
  str_split("\n") %>%
  unlist()

# create a matrix from lineages_string_by_row with a second var created from splitting by "\t"
lineages_mat <- lineages_string_by_row %>%
  str_split("\t", simplify = TRUE)

# convert lineages_mat to a data frame
lineages_df <- as.data.frame(lineages_mat)

# rename the var's "lineage_extracted", "description" and remove the first row
colnames(lineages_df) <- c("lineage_extracted", "description")
lineages_df <- lineages_df[,  c("lineage_extracted", "description")]

# remove row 1 that contains var names 
lineages_df_2 <- lineages_df[-1,] 

lineages_df_3 <- lineages_df_2 %>%
  filter(!(lineages_df_2$lineage_extracted == "" & lineages_df_2$description == ""))

lineages_df_4 <- lineages_df_3 %>%
  # create a 'status' var that is populated with either withdrawn or active
  mutate(status = case_when(
    # when 'withdrawn' is detected in the string then populate status with 'Withdrawn'
    str_detect(tolower(lineages_df_3$description), "withdrawn") ~ "Withdrawn",
    # otherwise populate status with 'Active'
    TRUE ~ "Active"
  ))

# extract string from beginning up until the first space or end of string and assign as lineage extracted. This steps is at times necessary due to errors in the file pulled. Within each row the lineage and descript should be separated by '\t' (see line 53). However at times a white space is mistakenly entered instead 
lineages_df_4$lineage_extracted <- str_extract(lineages_df_4$lineage_extracted, ".+?(?=$|[[:SPACE:]])")

# remove '*' from the lineage_extracted variables. Often withdrawn lineages are denoted with a '*' at the beginning
lineages_df_4$lineage_extracted <- str_remove_all(lineages_df_4$lineage_extracted, "\\*")

# assign lineages_df_4 to lineages_df_5
lineages_df_5 <- lineages_df_4
```

# If the pulled lineages file fails QA checks send an email stating so and stop the script.
```{r}
# check to make sure that all lineage_extracted follow a pattern of beginning one or more characters followed by zero or more combinations of a period followed by numbers. If there are any values that do not meet this check:
if(any(!str_detect(lineages_df_5$lineage_extracted, "^[[:upper:]]+(\\.[[:digit:]]+)*$"))) {
  
  # send an email stating there are values that do not follow the expected pattern
  email_from <- ""
  email_to <- ""
  email_subj <- "Sequencing - Lineages Output"
  email_body <- "There are values in the pulled Lineages file that are unexpected. Please check the pulled file and code."

  # send it
  sendmailR::sendmail(from = email_from,
                    to = email_to,
                    subject = email_subj,
                    msg = email_body,
                    headers= list("Reply-To" = email_from),
                    control = list(smtpServer = ""))

print("There are values in the pulled Lineages file that are unexpected. Script has been stopped and rendered up to this point.")  
}

# # running this line will give you all values that fail the check above
# lineages_df_5$lineage_extracted[!str_detect(lineages_df_5$lineage_extracted, "^[[:upper:]]+(\\.[[:digit:]]+)*$")]
```

# Exit run/render of script if there is a value in lineage_extracted  that does not follow a pattern of beginning one or more characters followed by zero or more combinations of a period followed by numbers. When this rmd is ran from a trigger script which uses render() the script will stop here and only render up to this point.
```{r}
#  exit if there are values in the pulled Lineages file that are unexpected
if(any(!str_detect(lineages_df_5$lineage_extracted, "^[[:upper:]]+(\\.[[:digit:]]+)*$"))) {
  knit_exit()
  stop("There are values in the pulled Lineages file that are unexpected. Script has been stopped and rendered up to this point.")
}
```

# Deduplicate based on repeated lineages. 
```{r}
# deduplicate any rows where lineage_extracted and status are the same (description is not a priority)
lineages_df_6 <- lineages_df_5 %>%
  distinct(lineage_extracted, status, .keep_all = TRUE)

# if there are any records where lineage_extracted is duplicated but status differs (when this occurs there should only ever be 2 records for each instances where lineage_extracted is duplicated; one with status == "Active" and one with status == "Withdrawn").
if (any(duplicated(lineages_df_6$lineage_extracted))) {
  
  # identify all records where lineage_extracted is duplicated 
  dup_records <- lineages_df_6[lineages_df_6$lineage_extracted %in% lineages_df_6$lineage_extracted[duplicated(lineages_df_6$lineage_extracted)],]
  
  # identify records where status == "Active"
  dup_records_active <- dup_records %>%
    filter(status == "Active")
  
  # remove records where lineage_extract is duplicated but status = "Active" for final output
  lineages_df_final <- anti_join(lineages_df_6, dup_records_active)
  
  # else if there no records with lineage_extracted is duplicated but status differs
} else {
  # lineages_df_6 is the final output
  lineages_df_final <- lineages_df_6
}

```

# Output and move files
```{r}
# move old file to archive


# output file

```

# Send an email when a new lineages file has been dropped into the Lineages folder.

