---
title: "sgtf_uw_pull"
author: "DIQA"
date: "12/20/2021"
output: html_document
---


# Load R libraries
```{r}
library(reticulate)
library(tidyverse)
library(fs)
library(lubridate)
library(here)
library(sendmailR)
library(knitr)
```

# Load python libraries
```{python}
import requests
from bs4 import BeautifulSoup
import csv
import re
```

# Read in required data_objects and define objects.
```{r}
# read in RDS containing vector of previously downloaded files
dir_prev_files <- readRDS("")

# define url
url = 'https://github.com/proychou/sgtf'
```

# get page content of landing page, pull correct/latest filename
```{python}
# get page contents
page = requests.get(r.url).text
# parse html
soup = BeautifulSoup(page, 'html.parser')
# find portion of html linking to available files
soup_find = soup.find_all('a', {'class': 'js-navigation-open Link--primary'} )
# extract names of available files
soup_file_names = [a.get_text() for a in soup_find]

# matches string containing 'county', ignores case
file_match = re.compile(".*county.*", re.IGNORECASE)
# return filename with 'county' somewhere in the name from soup_file_names
csv_filename = list(filter(file_match.match, soup_file_names))
#print the file name
print(csv_filename)

```

# If theres no new file send an email stating so and stop the script.
```{r}
# check if the csv_filename is in dir_prev_files (i.e this file was already previously downloaded). If there is no new file available send an email stating so.
if ((py$csv_filename %in% dir_prev_files) == TRUE) {
  email_from <- ""
  email_to <- ""
  email_subj <- "Sequencing - UW SGTF: No New Submission"
  email_body <- "No new UW SGTF file is available. Script has been stopped and rendered up to this point."

# send it
sendmailR::sendmail(from = email_from,
                    to = email_to,
                    subject = email_subj,
                    msg = email_body,
                    headers= list("Reply-To" = email_from),
                    control = list(smtpServer = ""))

print("No new UW SGTF file is available. Script has been stopped and rendered up to this point.")
}
```

# Exit run/render of script if csv_filename is detected in dir_prev_files. When this rmd is ran from a trigger script which uses render() the script will stop here and only render up to this point.
```{r}
# exit if the csv_filename is in dir_prev_files (i.e this file was already previously downloaded)
if ((py$csv_filename %in% dir_prev_files) == TRUE) {
  knit_exit()
}
```

# get page contents of sgtf file page, pull data
```{python}
# get page content
uw_sgtf_request = requests.get('https://raw.githubusercontent.com/proychou/sgtf/main/' + csv_filename[0])

# grab content of webpage
uw_sgtf_webcontent = uw_sgtf_request.content

# convert to string
uw_sgtf_webcontent_string = str(uw_sgtf_webcontent, 'utf-8')
```

# Transform data
```{r}
# call lineages_webcontent_string from python environment to R environment
uw_sgtf_webcontent_string <- py$uw_sgtf_webcontent_string

# split by "\n" to get a list by row, unlist
uw_sgtf_string_by_row <- uw_sgtf_webcontent_string %>%
  str_split("\n") %>%
  unlist

# create a matrix from lineages_string_by_row with a second var created from splitting by "\t"
sgtf_mat <- uw_sgtf_string_by_row %>%
  str_split(",", simplify = TRUE)

# convert lineages_mat to a data frame
sgtf_df <- as.data.frame(sgtf_mat)

# rename lineages_mat with vlues in first row
colnames(sgtf_df) <- dput(sgtf_mat[1,])

# remove row 1 that contains var names 
sgtf_df_1 <- sgtf_df[-1,] 

# exclude rows where all values are "" (blank)
sgtf_df_2 <- sgtf_df_1[rowSums(sgtf_df_1=="") != ncol(sgtf_df_1),]

# filter to where client_state == 'WA'
sgtf_df_3 <- sgtf_df_2 %>%
  filter(client_state == "WA")

# transform collect_date to as.date
sgtf_df_3$collect_date <-  as.Date(sgtf_df_3$collect_date)

# create 'WEEK_NUMBER' var containing week of year from 'collect_date', epiweek uses the US CDC epidemiological week and starts on Sun
sgtf_df_4 <- sgtf_df_3 %>%
  mutate('WEEK_NUMBER' = as.numeric(epiweek(sgtf_df_3$collect_date))) %>%
  mutate('n_all' = as.numeric(sgtf_df_3$n_all)) %>%
  mutate('n_sgtf' = as.numeric(sgtf_df_3$n_sgtf))

# create 'WEEK_START' var containing the start dates of the 'WEEK_NUMBER' intervals
sgtf_df_5 <- sgtf_df_4 %>%
  mutate('WEEK_START (MM/DD/YYYY)' = floor_date(as.Date(sgtf_df_4$collect_date), unit="week"))

# create 'WEEK_END' var containing the end dates of the 'WEEK_NUMBER' intervals
sgtf_df_6 <- sgtf_df_5 %>%
  mutate('WEEK_END (MM/DD/YYYY)' = ceiling_date(as.Date(sgtf_df_5$collect_date), unit="week") - 1)

sgtf_df_7 <- sgtf_df_6 %>%
  group_by(WEEK_NUMBER) %>%
  mutate(`TOTAL_POSITIVES (Total positives on an assay that can detect SGTF refer to Ct definitions below)` = sum(n_all)) %>%
  mutate('TOTAL_SGTF' = sum(n_sgtf)) %>%
  mutate('%_SGTF' = TOTAL_SGTF/`TOTAL_POSITIVES (Total positives on an assay that can detect SGTF refer to Ct definitions below)`)

sgtf_df_8 <- sgtf_df_7[, c(
  "WEEK_NUMBER", 
  "WEEK_START (MM/DD/YYYY)", 
  "WEEK_END (MM/DD/YYYY)", 
  "TOTAL_POSITIVES (Total positives on an assay that can detect SGTF refer to Ct definitions below)", 
  "TOTAL_SGTF", 
  "%_SGTF"
)]

sgtf_df_final <- distinct(sgtf_df_8)
```

# Check for previously output SGTF_UW_ file. If there is one then move it to the archive
```{r}
# get dir of sgtf submission folder
dir_sgtf_submissions <- dir_ls("",
       type = "file"
       )

# in dir_sgtf_submission if "SGTF_UW_.*.csv" is detected in file path index as TRUE, otherwise FALSE
dir_sgtf_submissions_uw_index <- str_detect(dir_sgtf_submissions, paste0("SGTF_UW_.*.csv"))

# for each index in dir_sgtf_submissions_uw_index
for (i in dir_sgtf_submissions_uw_index) {
  #if index == TRUE
  if (i == "TRUE") {
    # move the corresponding file in dir_sgtf_submission indexed by dir_sgtf_uw_index to the archive folder
    file_move(dir_sgtf_submissions[dir_sgtf_submissions_uw_index], 
              sub("(Submissions/)(SGTF_)", "\\1Archive/\\2", dir_sgtf_submissions[dir_sgtf_submissions_uw_index])
              )
  }
}

```

# Output file
```{r}
# output file
write_csv(sgtf_df_final, 
          file.path("", 
                    paste0("S-Dropouts/Submissions/SGTF_UW_", today(), ".csv")
                    ), 
          na = ""
          )
```

# If file is successfully output add csv_filename to dir_prev_files.rds and output an email. 
```{r}
# if SGTF_UW_today() exists in dir_sgtf_submissions
if (any(str_detect(dir_sgtf_submissions, paste0("SGTF_UW_", today(), ".csv")))) {
  
  new_dir_prev_file <- c(dir_prev_files, py$csv_filename)

  saveRDS(new_dir_prev_file, "")

  email_from <- ""
  email_to <- ""
  email_subj <- "Sequencing - UW SGTF: New Submission"
  email_body <- "There is a new UW SGTF file that has been dropped in the submissions folder."

  # send it
  sendmailR::sendmail(from = email_from,
                      to = email_to,
                      subject = email_subj,
                      msg = email_body,
                      headers= list("Reply-To" = email_from),
                      control = list(smtpServer = ""))
}
  
```
