---
title: "lab_submissions_monitoring"
author: "Nazmul Khan" 
date: "`r Sys.Date()`"
output: html_document
---

# Load R libraries.
```{r}
library(dplyr)
library(stringr)
library(lubridate)
library(writexl)
library(knitr)
library(tidyverse)
library(here)
library(openxlsx)
library(sendmailR)
library(glue)
```

# Tricks to make not-in operator
```{r}
`%!in%` <- Negate(`%in%`)
```

# Get the full list of all xlsx files from the root and sub folders in Completed Submissions
```{r}
# Get the full list of all xlsx files from the root folder and sub folders in the completed folders 
files<-list.files("", 
                  pattern = "\\.xlsx$|\\.csv$|\\.xls$", 
                  full.name=TRUE, 
                  recursive = TRUE
                  )

# remove file names with "_RUNNING.*
files<-files[!str_detect(files, "_RUNNING.*")]
```

# Capture folder names and the name/date of the most recent files in each folder
```{r}
# Get the corresponding folder names from the list of files in the completed folders
dirs <- dirname(files)

# get most recent file names from each of the completed folders
lastfiles <- tapply(files,dirs,function(v) v[which.max(file.mtime(v))])

# get the date of most recent file from each of the completed folders
mtime<-as.Date(file.info(lastfiles)$mtime,"%Y/%m/%d", tz = "EST5EDT")
```

# Bind recent dates and file names, create variable for folder name, label "Completed" folders
```{r}
# bind recent date to recent file names
recent_completed<-cbind(as.data.frame(lastfiles),as.data.frame(mtime))

#create a variable for folder name from the row names which captures the folder name
recent_completed$folders<-rownames(recent_completed)

# make a variable labeling completed folders
recent_completed$type<-"Completed"

# set the row names to null as they have been saved in the folder variable in step above 
rownames(recent_completed)=NULL
```

# Repeat above process for Submissions folder
```{r}
# get the full list of all xlsx files from the root folder and sub folders in the submitted folders 
files<-list.files("", 
                  pattern = "\\.xlsx$|\\.csv$|\\.xls$", 
                  full.name=TRUE, 
                  recursive = TRUE
                  )

# remove file names with "_RUNNING.*
files<-files[!str_detect(files, "_RUNNING.*")]

# get the corresponding folder names from the list of files in the submitted folders
dirs <- dirname(files)

# get most recent file names from each of the submitted folders
lastfiles <- tapply(files,dirs,function(v) v[which.max(file.mtime(v))])

# get the date of most recent file from each of the submitted folders
mtime<-as.Date(file.info(lastfiles)$mtime,"%Y/%m/%d", tz = "EST5EDT")

# bind recent date to recent file names
recent_submitted<-cbind(as.data.frame(lastfiles),as.data.frame(mtime))

# create a variable for folder name from the row names which captures the folder name
recent_submitted$folders<-rownames(recent_submitted)

# make a variable labeling submitted folders
recent_submitted$type<-"Submitted"

# set the row names to null as they have been saved in the folder variable in step above 
rownames(recent_submitted)=NULL
```

# Bind rows of completed and sumbitted
```{r}
recent_all<-rbind(recent_completed,recent_submitted)
```

# Get the difference in days from the last update to today's date 
```{r}
recent_all$datediff = as.numeric(difftime(Sys.Date(), recent_all$mtime, units = "days"))
```

# Clean up lab names
```{r}
recent_all<-recent_all %>%
  mutate(Lab=case_when( 
    grepl("/NW_Genomics/Altius_Crosswalk",folders) ~ "",
    grepl("Aegis",folders) ~ "Aegis",
    grepl("Altius",folders) ~ "Altius",
    grepl("Atlas_Genomics",folders) ~ "Atlas Genomics",
    grepl("Boise_VA",folders) ~ "Boise VA",
    grepl("CDC_Cumulative",folders) ~ "CDC Cumulative",
    grepl("Fulgent_Genetics",folders) ~ "Fulgent Genetics",
    grepl("Gravity_Diagnostics",folders) ~ "Gravity Diagnostics",
    grepl("Lauring_Lab",folders) ~ "Lauring Lab",
    grepl("NW_Genomics",folders) ~ "NW Genomics",
    grepl("Helix",folders) ~ "Helix",
    grepl("Infinity_Biologix",folders) ~ "Infinity Biologix",
    grepl("JacksonLab",folders) ~ "Jackson Lab",
    grepl("Kaiser",folders) ~ "Kaiser",
    grepl("Labcorp",folders) ~ "Labcorp",
    grepl("LabCorp",folders) ~ "Labcorp",
    grepl("OAMD",folders) ~ "OAMD",
    grepl("OHSU",folders) ~ "OHSU",
    grepl("PHL",folders) ~ "PHL",
    grepl("Providence_Swedish",folders) ~ "Providence Swedish",
    grepl("Quest",folders) ~ "Quest",
    grepl("SCAN_Bedford",folders) ~ "SCAN Bedford",
    grepl("USAFSAM",folders) ~ "US Air Force School of Aerospace",
    grepl("UW_Virology",folders) ~ "UW Virology",
    TRUE ~ "")) %>%
  select(Lab, everything()) %>%
  filter(Lab %!in% c("",
                     "Atlas Genomics",
                     "CDC Cumulative",
                     "Helix",
                     "Labcorp",
                     "OAMD",
                     "OHSU",
                     "PHL",
                     "Quest",
                     "SCAN Bedford"
  )) %>%
  arrange(Lab)

```

# Create a summary table with smallest datediff 
```{r}
summary_all<-recent_all%>%
  filter(Lab != "") %>%
  group_by(Lab) %>%
  filter(datediff==min(datediff)) %>%
  select(Lab, datediff)
```

# Clean up ties in datediff in summary table 
```{r}
summary_all<-distinct(summary_all)
```

# Sort summary table by lab and datediff
```{r}
# sort in descending order
summary_all<-summary_all %>%
  arrange(-datediff)

#sort detailed table by lab and datediff
recent_all<-recent_all %>%
  arrange(Lab,datediff)
```

# Set output Excel file name
```{r}
fname=paste0("",
             Sys.Date(),
             ".xlsx")
```

# Save tables to Excel file
```{r}
list_of_datasets <- list("Summary" = summary_all, "Detailed" = recent_all)

write.xlsx(list_of_datasets, file = fname, overwrite=T)
```

# Prepare email
```{r}

# Special message text
specialmessage <- glue::glue("*** This is an auto-generated email scheduled to run on Tuesday of every week. If there is additional information you would like added or modified, please contact the DIQA Genomic Sequencing team at ***")

# email body text
bdy <- glue::glue(
  "Hi all,


A weekly update of most recent lab files has been generated and saved in the folder 'Lab Monitoring' for your review.


{specialmessage} ")

```

# Send email
```{r}
length(recent_all)

if(nrow(recent_all)>0){
  email_from <- ""
  email_to <- c("")
  email_subj <- paste("Sequencing - Automated Email Lags in Receiving Lab Files on: ",Sys.Date())

  sendmailR::sendmail(email_from, 
                    email_to, 
                    email_subj,
                    msg = bdy,
                    headers= list("Reply-To" = email_from), control= list(smtpServer= ))}
```
