---
title: "QA_duplicates"
author: "DIQA"
date: "2/1/2022"
params: 
  test_mode: FALSE 
output: html_document
---

# Load Libraries
```{r}
library(DBI)
library(odbc)
library(lubridate)
library(tidyverse)
library(fs)
library(lubridate)
library(here)
library(sendmailR)
```

# Set the directory to either the Net Drive or local test envionment
```{r}
# By default set the project folder to the current R project folder
project_folder <- here()

# If running production mode, read from and write to the net drive folder
if(!params$test_mode) {
  project_folder <- readRDS(file.path(project_folder, "Data_Objects/dir_object.RDS"))
}

```

# Set up WDRS connection requirements
```{r}
conn_list <- readRDS(file.path(project_folder, "Data_Objects/conn_list.RDS"))

connection <- DBI::dbConnect(odbc::odbc(), 
                             Driver = conn_list[[1]], 
                             Server = conn_list[[2]], 
                             Database = conn_list[[3]], 
                             Trusted_connection = conn_list[[4]], 
                             ApplicationIntent = conn_list[[5]])
```

# Query ELR ENTIRE table for positive COVID labs
```{r}
#wdrs_entire <- dbGetQuery(connection, "
                      #SELECT DISTINCT 
                      #CASE_ID,
                      #FILLER__ORDER__NUM,
                      #SPECIMEN__COLLECTION__DTTM
                      #FROM [dbo].[DD_ELR_DD_ENTIRE]
                      #WHERE CODE = 'SARS' and WDRS__RESULT__SUMMARY!='Negative' 
                      #")
```

# Query COVID19 FLATTENED table for sequencing fields of interest
```{r, echo=FALSE}
wdrs_flattened <- dbGetQuery(connection, "
                    SELECT DISTINCT
                    CASE_ID,
                    FIRST_NAME,
                    LAST_NAME,
                    BIRTH_DATE,
                    CDC_N_COV_2019_SEQUENCE_ACCESSION_NUMBER,
                    CDC_N_COV_2019_SEQUENCE_CLINICAL_ACCESSION_NUMBER,
                    CDC_N_COV_2019_SEQUENCE_SPECIMEN_COLLECTION_DATE,
                    CDC_N_COV_2019_SEQUENCE_VARIANT_OPEN_TEXT
                    FROM [dbo].[DD_GCD_COVID_19_FLATTENED]
                    ")

# Separate delimited values into own row
wdrs_flat_clean <-wdrs_flattened %>% 
  separate_rows(CDC_N_COV_2019_SEQUENCE_ACCESSION_NUMBER,
                CDC_N_COV_2019_SEQUENCE_CLINICAL_ACCESSION_NUMBER, 
                CDC_N_COV_2019_SEQUENCE_VARIANT_OPEN_TEXT,
                CDC_N_COV_2019_SEQUENCE_SPECIMEN_COLLECTION_DATE, sep = ',') %>% 
  mutate(CDC_N_COV_2019_SEQUENCE_ACCESSION_NUMBER=str_trim(CDC_N_COV_2019_SEQUENCE_ACCESSION_NUMBER,side= "both")) %>%
  mutate(CDC_N_COV_2019_SEQUENCE_CLINICAL_ACCESSION_NUMBER=str_trim(CDC_N_COV_2019_SEQUENCE_CLINICAL_ACCESSION_NUMBER,side= "both")) %>%
  mutate(CDC_N_COV_2019_SEQUENCE_VARIANT_OPEN_TEXT=str_trim(CDC_N_COV_2019_SEQUENCE_VARIANT_OPEN_TEXT,side= "both")) %>%
  mutate(CDC_N_COV_2019_SEQUENCE_SPECIMEN_COLLECTION_DATE=str_trim(CDC_N_COV_2019_SEQUENCE_SPECIMEN_COLLECTION_DATE,side= "both"))
```

# Load WA GISAID data
```{r}
#wa_gisaid <- readRDS(file=file.path(project_folder,"GISAID Data/wa_gisaid.rds"))
```

# Read in files for review
```{r}

filenames <- list.files(file.path(project_folder, paste0("For_Review/to_process/")), full.names=TRUE, pattern="*.csv")

duplicates_template <- data.frame()

for (i in 1:length(filenames)){
  
temp_data <- read.csv(filenames[i], colClasses = c("SEQUENCE_CLINICAL_ACCESSION"="character"))
temp_data<-temp_data %>%
mutate(File_PATH=filenames[i])
duplicates_template <- bind_rows(duplicates_template, temp_data)
}

```

# Check if Sequence Accession is in WDRS (note this is the same as QA_SA_WDRS_DUPE but double checks the other SAs)
```{r}
# If the Sequence Accession is in WDRS, then we can ignore the record even if the lineage is different
# Molecular Epi has a process to look for significantly different changes in lineage, we do not need to keep track of these re submissions
# If the sequence accession is in WDRS, flag as "remove"

# Check if SA is already in WDRS
sa_wdrs_check <- wdrs_flat_clean %>% 
  filter(! is.na(CDC_N_COV_2019_SEQUENCE_ACCESSION_NUMBER)) %>%
  filter(CDC_N_COV_2019_SEQUENCE_ACCESSION_NUMBER !="") %>%
  inner_join(duplicates_template, by=c("CDC_N_COV_2019_SEQUENCE_ACCESSION_NUMBER" = "SEQUENCE_ACCESSION")) %>% 
  select(CDC_N_COV_2019_SEQUENCE_ACCESSION_NUMBER)

# Create remove flag
for_review <- duplicates_template %>% mutate(remove = ifelse(SEQUENCE_ACCESSION %in% sa_wdrs_check$CDC_N_COV_2019_SEQUENCE_ACCESSION_NUMBER, "1", "0")) #%>%
        
# Filter out non duplicates
for_review <- for_review %>% filter(!(is.na(QA_SA_INT_DUPE) & is.na(QA_SCA_INT_DUPE) & is.na(QA_SA_WDRS_DUPE) & is.na(QA_SCA_WDRS_DUPE)))

```

# Creating "count_yes" column to be consistent with other for_review scripts
```{r}
# Count the number of flags across a row for the QA flag column
for_review <- for_review %>%
  mutate(count_yes = rowSums(across(contains("QA_"), `%in%`, 1)))
```

# Some records were processed more than once and appear in for_review files multiple times with a different date in sequence note
# This should probably go in main script
```{r}
# Dedup on all relevant vars except sequence note
# Removing for now - this is only relevant when the same submissions are processed multiple times
#for_review <- distinct(for_review, CASE_ID, SEQUENCE_SGTF, SEQUENCE_SPECIMEN, SEQUENCE_REASON, SEQUENCE_DATE, 
                       #SEQUENCE_LAB, SEQUENCE_STATUS, SEQUENCE_REPOSITORY, SEQUENCE_ACCESSION,SEQUENCE_VARIANT_OPEN_TEXT, 
                       #SEQUENCE_CLINICAL_ACCESSION, SEQUENCE_SPECIMEN_COLLECTION_DATE, SEQUENCE_REVIEWED, .keep_all=TRUE)

```

# INTERNAL DUPLICATES
# Separate for_review flags - QA_SCA_INT_DUPE only, QA_SA_INT_DUPE only, or BOTH SCA/SA INT DUPE
# These are handled differently given the duplicate type
```{r}
# SCA is duplicated internally only
sca_int_dupe <- for_review %>% filter(QA_SCA_INT_DUPE==1 & is.na(QA_SA_INT_DUPE))

# SA duplicated internally only
sa_int_dupe <- for_review %>% filter(QA_SA_INT_DUPE==1 & is.na(QA_SCA_INT_DUPE))

# SCA and SA both duplicated internally
sca_sa_int_dupe <- for_review %>% filter(QA_SCA_INT_DUPE==1 & QA_SA_INT_DUPE==1)

```


# If SCA and SA are both duplicated internally, the two records likely match to two different case_ids
# Sometimes, one of the case_ids is clearly an incorrect match and is also marked as "QA_COLLECT_DATE"
# We can remove the bad match if it is also marked as QA_COLLECT_DATE
# Otherwise, flag as for manual review

```{r}

`%notin%` <- Negate(`%in%`)

# Filter for those SA/SCA pairs with a bad match based on collection date
sca_sa_int_collection_date <- sca_sa_int_dupe %>% group_by(SEQUENCE_ACCESSION, SEQUENCE_CLINICAL_ACCESSION, QA_COLLECT_DATE) %>% 
                              filter(n()==1) %>% 
                              mutate(remove = case_when((QA_COLLECT_DATE ==1) ~ 1, 
                                              (is.na(QA_COLLECT_DATE)) ~ 0)) %>% 
                              mutate(manual_review = case_when((QA_COLLECT_DATE==1) ~ 0, 
                                              (is.na(QA_COLLECT_DATE)) ~ 0))

# Filter for those SA/SCA pairs that match to one/the same case_id - this doesn't seem to happen frequently
sca_sa_int_same_case <- sca_sa_int_dupe %>% group_by(CASE_ID, SEQUENCE_ACCESSION, SEQUENCE_CLINICAL_ACCESSION) %>% filter(n()>1) %>% mutate(remove=0, manual_review =1) 

# Filter for those records that match to multiple/different case_ids regardless of collection date - send these to manual review
sca_sa_int_diff_case <- sca_sa_int_dupe %>% filter(CASE_ID %notin% sca_sa_int_collection_date$CASE_ID) %>% 
                        filter(CASE_ID %notin% sca_sa_int_same_case$CASE_ID) %>% mutate(remove =0, manual_review=1)


# Combine back into one dataframe 
sca_sa_int_dupe_reviewed <- rbind(sca_sa_int_diff_case, sca_sa_int_collection_date, sca_sa_int_same_case)

# Check that no records were lost
if (nrow(sca_sa_int_dupe) != (nrow(sca_sa_int_dupe_reviewed))) {
   print("Not all records are accounted for - locate missing records")
   } else{
   print("All SCA/SA internal duplicates reviewed and accounted for!")
   }

```

# SCA repeated internally only
# If only SCA is repeated internally, we can roster the record IF the case_ids are the same (SA must be different)

```{r}

# Group by SCA and CASE_ID - matches to the same case_id so good
same_sca_good <- sca_int_dupe %>% 
  group_by(CASE_ID, SEQUENCE_CLINICAL_ACCESSION) %>% 
  filter(n()>1) %>% mutate(remove =0, manual_review=0)

# Group by SCA and CASE_ID - matches to distinct case_ids so bad
same_sca_bad <- sca_int_dupe %>% 
  group_by(CASE_ID, SEQUENCE_CLINICAL_ACCESSION) %>% 
  filter(n()==1) %>% mutate(remove =0, manual_review=1)


# Combine back into one dataframe 
sca_int_dupe_reviewed <- rbind(same_sca_good, same_sca_bad)

```

# SA repeated internally only, SA is not in WDRS
# Often matches to two case_ids and therefore requires manual review

```{r}
# Flag these for manual review
sa_int_dupe_reviewed <- sa_int_dupe %>% mutate(remove =0, manual_review=1)
```

# Combine all the internal duplicates
```{r}
internal_duplicates_reviewed <- rbind(sca_sa_int_dupe_reviewed, sca_int_dupe_reviewed, sa_int_dupe_reviewed)
```

# WDRS DUPLICATES
# Separate for_review flags - QA_SCA_WDRS_DUPE only, QA_SA_WDRS_DUPE only, or BOTH SCA/SA WDRS DUPE
# These are handled differently given the duplicate type

```{r}
# SCA is duplicated with WDRS only
sca_wdrs_dupe <- for_review %>% filter(QA_SCA_WDRS_DUPE==1 & is.na(QA_SA_WDRS_DUPE))

# SA duplicated with WDRS only
sa_wdrs_dupe <- for_review %>% filter(QA_SA_WDRS_DUPE==1 & is.na(QA_SCA_WDRS_DUPE))

# SCA and SA both duplicated with WDRS
sca_sa_wdrs_dupe <- for_review %>% filter(QA_SCA_WDRS_DUPE==1 & QA_SA_WDRS_DUPE==1)

```

# SA WDRS duplicates only can be ignored
```{r}
sa_wdrs_dupe_reviewed <- sa_wdrs_dupe %>% mutate(remove =1, manual_review=0) 
```

# If SA and SCA are both duplicated with WDRS, they can be ignored as well
```{r}
sca_sa_wdrs_dupe_reviewed <- sca_sa_wdrs_dupe %>% mutate(remove =1, manual_review=0) 
```

# SCA repeated with WDRS only
# If only SCA is repeated with WDRS, we can roster the record IF the case_ids are the same (SA must be different)
```{r}
# Because SA is not in WDRS, we do not need to worry about creating a duplicate, we just need to make sure the case_ids are the same

# Find case_id associated with the SCA in WDRS
sca_wdrs_check <- wdrs_flat_clean %>% 
  filter(! is.na(CDC_N_COV_2019_SEQUENCE_CLINICAL_ACCESSION_NUMBER)) %>%
  filter(CDC_N_COV_2019_SEQUENCE_CLINICAL_ACCESSION_NUMBER !="") %>%
  inner_join(sca_wdrs_dupe, by=c("CDC_N_COV_2019_SEQUENCE_CLINICAL_ACCESSION_NUMBER" = "SEQUENCE_CLINICAL_ACCESSION")) %>% 
  select(CASE_ID.x, CASE_ID.y, CDC_N_COV_2019_SEQUENCE_CLINICAL_ACCESSION_NUMBER, CDC_N_COV_2019_SEQUENCE_ACCESSION_NUMBER, SEQUENCE_ACCESSION) %>%
  filter((CASE_ID.x==CASE_ID.y) & (CDC_N_COV_2019_SEQUENCE_ACCESSION_NUMBER!=SEQUENCE_ACCESSION)) #We want case_ids to be the same, but SA to be different

# Keep the SCAs that match to the same case_id and have different sequence accessions
sca_wdrs_dupe_good <- sca_wdrs_dupe %>%
                    filter(SEQUENCE_CLINICAL_ACCESSION %in% sca_wdrs_check$CDC_N_COV_2019_SEQUENCE_CLINICAL_ACCESSION_NUMBER) %>%
                    mutate(remove =0, manual_review=0)

`%notin%` <- Negate(`%in%`)
sca_wdrs_dupe_bad <- sca_wdrs_dupe %>%
                filter(SEQUENCE_CLINICAL_ACCESSION %notin% sca_wdrs_check$CDC_N_COV_2019_SEQUENCE_CLINICAL_ACCESSION_NUMBER) %>%                     
                  mutate(remove=0, manual_review=1)

sca_wdrs_dupe_reviewed <- rbind(sca_wdrs_dupe_good, sca_wdrs_dupe_bad)

```

# Combine all the WDRS duplicates
```{r}
wdrs_duplicates_reviewed <- rbind(sca_wdrs_dupe_reviewed, sca_sa_wdrs_dupe_reviewed, sa_wdrs_dupe_reviewed)
```

# Combine back together internal and wdrs duplicates
```{r}
reviewed <- rbind(wdrs_duplicates_reviewed, internal_duplicates_reviewed)

# Check that no records were lost
if (nrow(reviewed) != (nrow(for_review))) {
   print("Not all records are accounted for - locate missing records")
   } else{
   print("All internal and WDRS duplicates reviewed!")
   }

```
# Separate out by result type - roster_prelim, manual review, or remove
```{r}
remove <- reviewed %>% filter(remove==1)

manual_review <- reviewed %>% filter(manual_review==1)

roster_prelim <- reviewed %>% filter(remove==0 & manual_review==0)

# Change QA duplicate columns blank and recalculate count_yes
# If record still needs review for other non-duplicate reasons, it will be sent back to for_review
roster_prelim <- roster_prelim %>% mutate_at(vars(c("QA_SCA_INT_DUPE", "QA_SA_INT_DUPE", "QA_SCA_WDRS_DUPE", "QA_SA_WDRS_DUPE")), ~na_if(., 1))

# Count the number of flags across a row for the QA flag column
roster_prelim <- roster_prelim %>%
  mutate(count_yes = rowSums(across(contains("QA_"), `%in%`, 1)))

```


# If a record still has other quality issues, send back to for_review
```{r}
# If count_yes >0 then there are other quality issues present and must be sent back
roster_quality_issues <- roster_prelim %>% filter(count_yes>0)

if (nrow(roster_quality_issues>0)) {
   print("Some records still have quality issues")
  
  write.csv(roster_quality_issues, file.path(project_folder, paste0("For_Review/to_process/duplicates_other_QA.csv")), na="",row.names = FALSE)
  
   } else{
   print("No additional QA checks")
   }

# If count_yes==0 then record has not other quality issues and can be rostered!
roster_final <- roster_prelim %>% filter(remove==0 & manual_review==0 & count_yes==0)

```

# Prepare Roster and output to write_roster_here
```{r}

roster_final <- roster_final %>% select (
          "CASE_ID",
          "SEQUENCE_SGTF",
          "SEQUENCE_SPECIMEN",
          "SEQUENCE_DATE",
          "SEQUENCE_REASON",
          "SEQUENCE_LAB",
          "SEQUENCE_STATUS",
          "SEQUENCE_REPOSITORY",
          "SEQUENCE_ACCESSION",
          #"SEQUENCE_EPI_ISL",
          "SEQUENCE_VARIANT_OPEN_TEXT",
          "SEQUENCE_CLINICAL_ACCESSION", 
          "SEQUENCE_SPECIMEN_COLLECTION_DATE",
          #"SEQUENCE_ROSTER_PREPARE_DATE",
          "SEQUENCE_NOTES",
          "SEQUENCE_REVIEWED",
          "Case.Note")


if (nrow(roster_final>0)) {
   print("Records sent to write_roster_here")
  
  write.csv(roster_final, file.path(project_folder, paste0("write_roster_here/duplicates_roster_",today(),".csv")), na="",row.names = FALSE)

   } else{
   print("No records for roster today")
   }

```

# Output duplicate records that require manual review from DIQA/Data Support
```{r}
# Write records with manual review issues to folder for DIQA/Data Support review

if (nrow(roster_final>0)) {
   print("Records sent to manual review")
  
  write.csv(manual_review, file.path(project_folder, paste0("For_Review/DIQA Test/duplicates_manual_review_",today(),".csv")), na="",row.names = FALSE)

   } else{
   print("No records for manual review today")
   }

```

# Send email - Will update once processes are established
```{r}
email_from <- ""
email_to <- c("")
email_subject <- "Sequencing - Duplicates Reviewed"
email_body <- "Duplicate records processed today"

if(!params$test_mode) {
  sendmailR::sendmail(from = email_from,
                      to = email_to,
                      subject = email_subject,
                      msg = email_body,
                      headers= list("Reply-To" = email_from),
                      control = list(smtpServer = "")
                      )
}

```



