---
title: "Fuzzy Collection Date Protocol"
author: "Frank Aragona"
date: "12/29/2021"
output:
  rmdformats::readthedown:
    highlight: espresso
    code_folding: hide
params: 
  test_mode: FALSE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

This is sequencing data from a submitter that was fuzzy matched on name and DOB but could not be matched to the specimen collection date for an event in WDRS. We currently have a 14 day window in which a collection date can match to a test in WDRS. In other words, this data has sequencing information from a date that could not be match to a specimen date for an individual in WDRS. **This ignores data that was FAILED/LOW QUALITY** 

This protocol is broken down into 2 main parts, each with guidance on what to do with the data:

1.  Matched - Sequencing data that could be matched to WDRS with no errors other than the collection date
2.  Unmatched - Sequencing data that could NOT be matched and may have multiple errors

There is a xlsx file in `For_Review/Fuzzy_matches/Combined` that has all the combined data and a sheet for matched and unmatched to make processing the data easier.

For transparency, you can click on the `Code` buttons to show all my code and how data was derived

```{r,warning=FALSE,message=FALSE}
library(DBI)
library(odbc)
library(lubridate)
library(tidyverse)
library(readxl)
library(here)
library(fs)
library(sendmailR)
library(fuzzyjoin)
library(kableExtra)
# library(miniUI)
```


```{r,warning=FALSE,message=FALSE}
# Read in r_creds.rds
r_creds <-readRDS(file.path(Sys.getenv("USERPROFILE"), "Projects/Sequencing/Data_Objects", "r_creds.RDS"))

# Set up connection to WDRS
connection <- DBI::dbConnect(odbc::odbc(), 
                             Driver = r_creds$conn_list[1], 
                             Server = r_creds$conn_list[2], 
                             Database = r_creds$conn_list[3], 
                             Trusted_connection = r_creds$conn_list[4], 
                             ApplicationIntent = r_creds$conn_list[5])
```


```{r,warning=FALSE,message=FALSE}
#Compile all the csv with date mismatch in For_Review folder

filenames <- list.files(paste0(""),pattern="*csv",full.names=TRUE)

combine_fuzzy_date <- data.frame()

for (i in 1:length(filenames)){
  temp_data <- read.csv(filenames[i], colClasses = c("SEQUENCE_CLINICAL_ACCESSION"="character"))
  temp_data<-temp_data %>%
  mutate(File_PATH=filenames[i])
  combine_fuzzy_date <- bind_rows(combine_fuzzy_date, temp_data)
}

```


```{r}
# Move files into archive folder with today's date

if(length(filenames > 0)){

  # Create folder in Archive
  dir_create(paste0(file.path(""), today()))
   
  # Move files to archive
  file_move(filenames, paste0(file.path(""), today()))
}

```

```{r,warning=FALSE,message=FALSE}
#Clean the final csv
all_combined<- distinct(combine_fuzzy_date,across(!SEQUENCE_NOTES)) %>%
  as_tibble()
```


```{r}
# Separate out QA_MULTIPLE_MATCH - manual review for now
duplicates <- all_combined %>% filter(!(is.na(QA_MULTIPLE_MATCH)))

# Separate out QA_COLLECT_DATE
date_mismatch <- all_combined %>% filter(QA_COLLECT_DATE==1 & is.na(QA_MULTIPLE_MATCH))

# Some records are missing a flag, rare but happens - manual review for now
other <- all_combined %>% filter(is.na(QA_MULTIPLE_MATCH) & is.na(QA_COLLECT_DATE))

```


```{r}
# Send Multiple Matches and "other" records to manual review
if (nrow(duplicates>0)) {
   print("Duplicate Records sent to manual review")
  
   write.csv(duplicates, file.path(paste0("",today(),".csv")), na="",row.names = FALSE)

   } else{
   print("No duplicate records for review today")
   }


# There shouldn't be any records without a flag anymore but just in case
if (nrow(other>0)) {
   print("Other Records sent to manual review")
  
   write.csv(other, file.path(paste0("",today(),".csv")), na="",row.names = FALSE)

   } else{
   print("No other records for review today")
   }

```


```{r,warning=FALSE,message=FALSE}
#Connect to WDRS flattened table
wdrs_Flat <- dbGetQuery(connection, "
                    SELECT DISTINCT CASE_ID, CDC_N_COV_2019_SEQUENCE_ACCESSION_NUMBER,
                    CDC_N_COV_2019_SEQUENCE_CLINICAL_ACCESSION_NUMBER, CDC_N_COV_2019_SEQUENCE_SPECIMEN_COLLECTION_DATE
                    FROM [dbo].[DD_GCD_COVID_19_FLATTENED]
                    ")
```


```{r,warning=FALSE,message=FALSE}
#Clean the flat table
wdrs_Flat_Clean <-wdrs_Flat %>% 
  rename("SEQUENCE_ACCESSION"="CDC_N_COV_2019_SEQUENCE_ACCESSION_NUMBER") %>% 
  rename("SEQUENCE_CLINICAL_ACCESSION"="CDC_N_COV_2019_SEQUENCE_CLINICAL_ACCESSION_NUMBER") %>% 
  rename("SEQUENCE_SPECIMEN_COLLECTION_DATE"="CDC_N_COV_2019_SEQUENCE_SPECIMEN_COLLECTION_DATE") %>%
  separate_rows(SEQUENCE_ACCESSION,SEQUENCE_CLINICAL_ACCESSION, sep = ',') %>% 
  mutate(SEQUENCE_ACCESSION=str_trim(SEQUENCE_ACCESSION,side= "both")) %>%
  mutate(SEQUENCE_CLINICAL_ACCESSION=str_trim(SEQUENCE_CLINICAL_ACCESSION,side= "both"))
```

```{r,warning=FALSE,message=FALSE}
#Connect to the WDRS entire table
wdrs_ent <- dbGetQuery(
  connection,
  "
  SELECT Distinct CASE_ID,
    [FILLER__ORDER__NUM] as SEQUENCE_CLINICAL_ACCESSION,
    SPECIMEN__COLLECTION__DTTM as COLLECTION_DATE
  FROM [dbo].[DD_ELR_DD_ENTIRE]
  WHERE CODE = 'SARS'
  "
)
```


```{r,warning=FALSE,message=FALSE}
#Connect to GISAID in order to double check if the sequence exists in GISAID (it should)
wa_gisaid <- read_rds("")

names(wa_gisaid)<-str_replace_all(names(wa_gisaid), c(" " = "." , "," = "" ))

wa_gisaid2 <- wa_gisaid %>%
  mutate(SEQUENCE_ACCESSION=Virus.name)
```


```{r,warning=FALSE,message=FALSE}
#NoMatch is all data with no sequencing event in WDRS for that specific sequence accession number
NoMatchToWDRS<-anti_join(date_mismatch, wdrs_Flat_Clean, by="SEQUENCE_ACCESSION") %>%
  mutate(CASE_ID = as.character(CASE_ID))

#Match is data that does have a sequencing event for that sequence accession in wdrs
MatchToWDRS <- inner_join(date_mismatch,wdrs_Flat_Clean,by=c("SEQUENCE_ACCESSION","SEQUENCE_SPECIMEN_COLLECTION_DATE"))
```

# Match to GISAID
Try to pull the SEQUENCE_SPECIMEN_COLLECTION_DATE if it's NULL in the fuzzy output.

```{r}
gisaid_match <- left_join(date_mismatch,wa_gisaid2,by="SEQUENCE_ACCESSION")
```


# Matched to exisiting Sequence Event

This is data that could either be:

1.    Matched to a sequencing event already in WDRS except the collection date is outside the 14 day window.
2.    Need to be taken out because it is FAILED/LOW QUALITY

_How to Fix:_

*   Delete these rows in the csv as they already have the sequencing event or they are FAILED/Low Quality

Please see the file in `For_Review/Fuzzy_matches/Combined` under the `Matched` tab for this data. 

```{r,echo=F,eval=T}
to_delete <- bind_rows(MatchToWDRS,
                         date_mismatch[which(date_mismatch$SEQUENCE_STATUS =="FAILED" | date_mismatch$SEQUENCE_STATUS == "LOW QUALITY" ),])

if(nrow(to_delete)==0){
  print("There is currently no data that could be matched to a sequencing event already in WDRS")
}else{
  print("Please see the 'QAfuzzy_date_mismatch' csv for the matched data")
}
```


```{r,echo=T,eval=F}
#This chunk is only to show the code in the RMD output

#Match is data that does have a sequencing event for that sequence accession in wdrs
MatchToWDRS <- inner_join(date_mismatch,wdrs_Flat_Clean,by=c("SEQUENCE_ACCESSION","SEQUENCE_SPECIMEN_COLLECTION_DATE"))
 
to_delete <- bind_rows(MatchToWDRS,
                        date_mismatch[which(date_mismatch$SEQUENCE_STATUS =="FAILED" | date_mismatch$SEQUENCE_STATUS == "LOW QUALITY" ),])%>%
  select(CASE_ID,2:19)
# MatchToWDRS %>%
#   kbl(caption = "List of CASE_ID that have a sequencing event in WDRS and can be rostered")%>%
#   kable_styling()
```


# Unmatched - No exisiting Sequencing Event

This is data that could not be matched to a sequencing event in WDRS. There are 3 different ways to categorize this data, each with different ways to fix them.

1.    Data that can be matched to a positive event within 30 days AND match to a SCA for that event
2.    Data that can be matched to a positive event within 30 days with NO match to a SCA for that event
3.    Data that could not be matched to either a positive event within 30 days or a SCA

```{r,warning=FALSE,message=FALSE}
#NoMatch is all data with no sequencing event in WDRS for that specific sequence accession number
NoMatchToWDRS<-anti_join(date_mismatch, wdrs_Flat_Clean, by="SEQUENCE_ACCESSION") %>%
  mutate(CASE_ID = as.character(CASE_ID))
#join the unmatched to GISAID to see if there's a sequence
FoundSeq <-
  left_join(NoMatchToWDRS, wa_gisaid2, by = "SEQUENCE_ACCESSION") %>%
  mutate(CASE_ID=as.character(CASE_ID))%>%
  mutate(COLLECTION_DATE_WDRS = as.Date(mdy(COLLECTION_DATE_WDRS),format = "%Y-%m-%d"))%>%
  mutate(SEQUENCE_SPECIMEN_COLLECTION_DATE = as.Date(mdy(SEQUENCE_SPECIMEN_COLLECTION_DATE),format = "%Y-%m-%d"))%>%
  select(-Collection.date)

#prep data for date fuzzy match
wdrs_ent$match_date_minus3 <- as.Date(wdrs_ent$COLLECTION_DATE - 30)
wdrs_ent$match_date_plus3 <- as.Date(wdrs_ent$COLLECTION_DATE + 30)
wdrs_ent$SCA <- wdrs_ent$SEQUENCE_CLINICAL_ACCESSION

# First match the ones in entire to make the process faster
entire_table<- inner_join(wdrs_ent,FoundSeq,by = "CASE_ID")%>%
  select(CASE_ID,SCA,match_date_minus3,match_date_plus3,COLLECTION_DATE)%>%
  distinct()
```

## 1.   Match to Positive Event <= 30 days and a SCA

_Details:_

Here I'm taking the data that was unmatched to WDRS flat (but matched to GISAID) and trying to find a positive test for that CASE_ID, SCA, and date in the entire table. **In other words**, are there any positive tests for that person/SCA/date that match the sequencing data?

The sequencing data can be attached to these events regardless of specimen collection date (unless we see other issues?)

_How to Fix:_

*   There are some records here that could match to multiple events based on collection date/SCA. If that is the case **Assign the seq info to the event that has the closest collection date to match it and the correct SCA**
*   For the rest, **Assign the seq info to the correct event date (the single event) with the correct SCA**

```{r,warning=FALSE}
#Take the unmatched to WDRS flat (but matched to GISAID) and try to find a positive test for that CASE_ID, SCA, and date in the flattened table
fuzzy_date_SCA<-fuzzy_inner_join(FoundSeq, entire_table,
                                 by = c("CASE_ID" = "CASE_ID",
                                        "SEQUENCE_CLINICAL_ACCESSION" = "SCA",
                                        'SEQUENCE_SPECIMEN_COLLECTION_DATE' = "match_date_minus3",
                                        'SEQUENCE_SPECIMEN_COLLECTION_DATE' = "match_date_plus3"),
                                 match_fun = list(`==`, `==`, `>`, `<`))%>%
  # select(CASE_ID=CASE_ID.x,NAME_WDRS,SEQUENCE_CLINICAL_ACCESSION,
  #        SEQUENCE_SPECIMEN_COLLECTION_DATE,Date_Ent_Table=COLLECTION_DATE,SCA_Ent_Table=SCA,
  #        SEQUENCE_ACCESSION,SEQUENCE_VARIANT_OPEN_TEXT)%>%
  select(CASE_ID = CASE_ID.x,2:23,Date_Ent_Table=COLLECTION_DATE,SCA_Ent_Table=SCA) %>%
  distinct()

  
```

```{r,warning=FALSE}
if(nrow(fuzzy_date_SCA)!=0){
# These match on date, AND SCA. 
library(gt)
fuzzy_table <- fuzzy_date_SCA %>%
  gt(rowname_col ="NAME_WDRS")

ft2 <- fuzzy_table %>% 
  tab_spanner(
    label = "Sequence Info",
    columns = c(NAME_WDRS,SEQUENCE_CLINICAL_ACCESSION,
                   SEQUENCE_SPECIMEN_COLLECTION_DATE,SEQUENCE_ACCESSION,SEQUENCE_VARIANT_OPEN_TEXT)
  ) %>%
  tab_spanner(
    label = "Matched Event in Entire Table",
    columns = c(COLLECTION_DATE_WDRS,SCA_Ent_Table)
  )

ft3 <- ft2 %>%
  tab_style(
    style = list(
      cell_fill(color = "lightblue"),
      cell_text(weight = "bold")
    ),
    locations = cells_body(
      columns = c(Date_Ent_Table,SCA_Ent_Table,SEQUENCE_CLINICAL_ACCESSION,
                  SEQUENCE_SPECIMEN_COLLECTION_DATE)
    )) 
# %>%
#   cols_width(
#     c(Date_Ent_Table) ~ px(2000),
#   )


ft4<- ft3 %>%
  tab_style(style = list(cell_text(font = "Oswald",
                                   align = "left")),
            locations = list(cells_body(
              columns = c(everything())
            ))) %>%
  cols_label(
    SEQUENCE_CLINICAL_ACCESSION = "SEQUENCE\n CLINICAL_ACCESSION",
    SEQUENCE_SPECIMEN_COLLECTION_DATE = "SEQUENCE_SPECIMEN\n COLLECTION_DATE",
    SEQUENCE_ACCESSION = "SEQUENCE\n ACCESSION",
    SEQUENCE_VARIANT_OPEN_TEXT = "SEQUENCE\n VARIANT",
    Date_Ent_Table = "Collection Date",
    SCA_Ent_Table = "SEQUENCE\n CLINICAL_ACCESSION"
  )


ft5<- ft4 %>%
  tab_header(
    title = md("List of CASE_ID that match SCA and have <= 30 day collection date match"),
  subtitle = "Compares Sequencing info from submitters to matched event in WDRS_Entire") %>%
  tab_options(
    heading.background.color = "#c6dbef",
    column_labels.background.color = "#edf8fb",
    table.width = "60%"
    
  )

ft5 
} else{
  print("No data could be matched to a SCA/positive event within 30 days")
}
```

## 2.   Match to Positive Event <= 30 days but not a SCA

_Details:_

These cases have a positive test that matches the date of the sequencing information, but the SCA does not match (or doesn't exist for those tests)

**However!!** They have may have other tests in WDRS that match on SCA but don't match on the date. Therefore, I believe the SCA is not correct in WDRS and should actually be assigned to the test that has the correct sequencing date. In other words:

1.  The sequencing info here matches on date
2.  The SCA is attached to the wrong test (it should be attached to the test with the correct sequencing date)

_How to Fix:_ 

*   **For these records, attach the sequencing information to the positive event with the correct date**
*   Or if you think it would be best to reassign the SCA to the correct positive test, then do that

```{r,warning=FALSE}
#These could be matched on date but not on SCA
fuzzy_noSCA_<-fuzzy_inner_join(FoundSeq, entire_table,
                               by = c("CASE_ID" = "CASE_ID",
                                      'SEQUENCE_SPECIMEN_COLLECTION_DATE' = "match_date_minus3",
                                      'SEQUENCE_SPECIMEN_COLLECTION_DATE' = "match_date_plus3"),
                               match_fun = list(`==`, `>`, `<`))%>%
  # select(CASE_ID=CASE_ID.x,NAME_WDRS,SEQUENCE_CLINICAL_ACCESSION,
  #        SEQUENCE_SPECIMEN_COLLECTION_DATE,Date_Ent_Table=COLLECTION_DATE,SCA_Ent_Table=SCA,
  #        SEQUENCE_ACCESSION,SEQUENCE_VARIANT_OPEN_TEXT)%>%
  select(CASE_ID = CASE_ID.x,2:23,Date_Ent_Table=COLLECTION_DATE,SCA_Ent_Table=SCA) %>%
  distinct()

#exclude the ones that we already matched by SCA and date
fuzzy_noSCA <- anti_join(fuzzy_noSCA_,fuzzy_date_SCA,by = c("CASE_ID","NAME_WDRS"))

# if(nrow(fuzzy_noSCA)!=0){
# fuzzy_noSCA %>%
#   kbl(caption = "List of CASE_ID that have a positive test match but a SCA attached to the wrong test")%>%
#   kable_styling()
# }else{
#   print("No Records found - Move on!")
# }


```

```{r,warning=FALSE}
if(nrow(fuzzy_noSCA)!=0){
library(gt)
fuzzy_table2 <- fuzzy_noSCA %>%
  gt(rowname_col ="NAME_WDRS")

ft2 <- fuzzy_table2 %>% 
  tab_spanner(
    label = "Sequence Info",
    columns = c(NAME_WDRS,SEQUENCE_CLINICAL_ACCESSION,
                   SEQUENCE_SPECIMEN_COLLECTION_DATE,SEQUENCE_ACCESSION,SEQUENCE_VARIANT_OPEN_TEXT)
  ) %>%
  tab_spanner(
    label = "Entire Table",
    columns = c(Date_Ent_Table,SCA_Ent_Table)
  )

ft3 <- ft2 %>%
  tab_style(
    style = list(
      cell_fill(color = "lightblue"),
      cell_text(weight = "bold")
    ),
    locations = cells_body(
      columns = c(Date_Ent_Table,SCA_Ent_Table,SEQUENCE_CLINICAL_ACCESSION,
                  SEQUENCE_SPECIMEN_COLLECTION_DATE)
    ))

ft4<- ft3 %>%
  tab_style(style = list(cell_text(font = "Oswald",
                                   align = "left")),
            locations = list(cells_body(
              columns = c(everything())
            ))) %>%
  cols_label(
    SEQUENCE_CLINICAL_ACCESSION = "SEQUENCE\n CLINICAL_ACCESSION",
    SEQUENCE_SPECIMEN_COLLECTION_DATE = "SEQUENCE_SPECIMEN\n COLLECTION_DATE",
    SEQUENCE_ACCESSION = "SEQUENCE\n ACCESSION",
    SEQUENCE_VARIANT_OPEN_TEXT = "SEQUENCE\n VARIANT",
    Date_Ent_Table = "Collection Date",
    SCA_Ent_Table = "SEQUENCE\n CLINICAL_ACCESSION"
  )


ft5<- ft4 %>%
  tab_header(
    title = md("List of CASE_ID that match <= 30 day collection date match BUT don't match a SCA"),
  subtitle = "Compares Sequencing info from submitters to matched event in WDRS_Entire")%>%
  tab_options(
    heading.background.color = "#c6dbef",
    column_labels.background.color = "#edf8fb",
    table.width = "60%"
    
  )


ft5
}else{
  print("No Records found - Move on!")
}

```

## 3.   Data with No Match to positive test date or SCA

_Details:_

For these ones, we will have to create a new event in WDRS.

These records did not match because they either:

1.    Don't have a positive test in WDRS or, 
2.    May have a positive test but it's out of the date range of the new seq date 

_How to Fix:_

*   **Either way, we need to create a new seq field with that seq information, even though it's not attached to a positive test**
*   Or we could reach out to labs if you think that is necessary. 

```{r,warning=FALSE}
addressed_data<- bind_rows(fuzzy_date_SCA,fuzzy_noSCA) %>%
  distinct(CASE_ID)

  
stillnomatch<-anti_join(NoMatchToWDRS,addressed_data,by="CASE_ID") 

# Take out records that may have different case Id's for the same person as they will appear in 2 different tabs
stillnomatch_dedup <- stillnomatch %>%
  filter(!stillnomatch$NAME_WDRS %in% fuzzy_date_SCA$NAME_WDRS & !stillnomatch$SEQUENCE_ACCESSION %in% fuzzy_date_SCA$SEQUENCE_ACCESSION)



# if(nrow(stillnomatch)!=0){
# stillnomatch %>%
#   kbl(caption = "List of CASE_ID that need a new sequencing event")%>%
#   kable_styling()
# }else{
#   print("No records found - You're done! Go take a break:)")
# }

```

```{r,warning=FALSE}
if(nrow(stillnomatch_dedup)!=0){
  fuzzy_table3 <- stillnomatch_dedup %>%
    select(CASE_ID,NAME_WDRS,SEQUENCE_ACCESSION, SEQUENCE_CLINICAL_ACCESSION,SEQUENCE_VARIANT_OPEN_TEXT,SEQUENCE_SPECIMEN_COLLECTION_DATE,COLLECTION_DATE_WDRS)%>%
    gt(rowname_col ="NAME_WDRS")
  
  ft2 <- fuzzy_table3 %>% 
    tab_spanner(
      label = "Sequence Info",
      columns = c(NAME_WDRS,SEQUENCE_CLINICAL_ACCESSION,
                  SEQUENCE_SPECIMEN_COLLECTION_DATE,SEQUENCE_ACCESSION,SEQUENCE_VARIANT_OPEN_TEXT)
    )%>%
    tab_spanner(
      label = "Entire Table",
      columns = c(COLLECTION_DATE_WDRS)
    ) 
  
  
  ft3<- ft2 %>%
    tab_style(style = list(cell_text(font = "Oswald",
                                     align = "left")),
              locations = list(cells_body(
                columns = c(everything())
              ))) %>%
    cols_label(
      SEQUENCE_CLINICAL_ACCESSION = "SEQUENCE\n CLINICAL_ACCESSION",
      SEQUENCE_SPECIMEN_COLLECTION_DATE = "SEQUENCE_SPECIMEN\n COLLECTION_DATE",
      SEQUENCE_ACCESSION = "SEQUENCE\n ACCESSION",
      SEQUENCE_VARIANT_OPEN_TEXT = "SEQUENCE\n VARIANT",
      COLLECTION_DATE_WDRS = "COLLECTION DATE\n WDRS"
    )
  
  
  ft4<- ft3 %>%
    tab_header(
      title = md("List of CASE_ID that do not match a SCA or an event date"),
      subtitle = "Compares Sequencing info from submitters to matched event in WDRS_Entire")%>%
  tab_options(
    heading.background.color = "#c6dbef",
    column_labels.background.color = "#edf8fb",
    table.width = "60%"
    
  )

  
  ft4
}else{
  print("No records found - You're done! Go take a break:)")
}
```


```{r,warning=FALSE}
# Combine rough matches as these already have appropriate events in WDRS and will both need a new sequencing field as a fix. Same fix, different match

create_new_seq<- bind_rows(fuzzy_date_SCA,fuzzy_noSCA)
  

#Create an excel file with all the combined data if there are both unmatch and matched
if(nrow(NoMatchToWDRS)!=0){
  library(writexl)
sheets <- list("All Data Combined" = date_mismatch, "Delete these rows" = to_delete, "Create New Seq Field" = create_new_seq, "Create New Event" = stillnomatch_dedup) 
write_xlsx(sheets, paste0("",
    "fuzzy_date_mismatches_",
    today(), ".xlsx"
  ))
}else{
  print("There's no data OR somethin' ain't right")
}
```
