---
title: "MissingLineagesReview"
author: "Marcela/DIQA"
date: "9/14/2021"
params: 
  test_mode: FALSE
output: html_document
---
### Make note of which labs have missing CASE_IDs if it happens again - then we can update those scripts in the submitting scripts
```{r setup, include=FALSE}
knitr::opts_knit$set(echo = TRUE)
```

```{r}
library(DBI)
library(odbc)
library(lubridate)
library(tidyverse)
library(readxl)
library(here)
library(fs)
library(sendmailR)
# library(shiny)
# library(miniUI)
```

# Set the directory to either the Net Drive or local test envionment.
```{r}
# By default set the project folder to the current R project folder
project_folder <- here()

# If running production mode, read from and write to the net drive folder
if(!params$test_mode) {
  project_folder <- readRDS(paste0(project_folder, "/Data_Objects/dir_object.RDS"))
}

```

#Read in email lists
```{r}
email_list1 <- readRDS(file.path(project_folder, "Data_Objects/for_review/matchedlin_email.rds"))
email_list2 <- readRDS(file.path(project_folder, "Data_Objects/for_review/notfound_email.rds"))
```

Connecting to WDRS
```{r, echo=FALSE}
conn_list <- readRDS(paste0(project_folder,"/Data_Objects/conn_list.RDS"))


connection <- DBI::dbConnect(odbc::odbc(),
                             Driver = conn_list[[1]],
                             Server = conn_list[[2]],
                             Database = conn_list[[3]],
                             Trusted_connection = conn_list[[4]],
                             ApplicationIntent = conn_list[[5]])


```

Pulling data from the Flattened table
```{r, echo=FALSE}
wdrs_Flat <- dbGetQuery(connection, "
                    SELECT DISTINCT CDC_N_COV_2019_SEQUENCE_ACCESSION_NUMBER,
                    CDC_N_COV_2019_SEQUENCE_CLINICAL_ACCESSION_NUMBER
                    FROM [dbo].[DD_GCD_COVID_19_FLATTENED]
                    ")
```

Prepping the data from the Flattened table
```{r, echo=FALSE}

wdrs_Flat_Clean <-wdrs_Flat %>% 
  rename("SEQUENCE_ACCESSION"="CDC_N_COV_2019_SEQUENCE_ACCESSION_NUMBER") %>% 
  rename("SEQUENCE_CLINICAL_ACCESSION"="CDC_N_COV_2019_SEQUENCE_CLINICAL_ACCESSION_NUMBER") %>% 
  separate_rows(SEQUENCE_ACCESSION,SEQUENCE_CLINICAL_ACCESSION, sep = ',') %>% 
  mutate(SEQUENCE_ACCESSION=str_trim(SEQUENCE_ACCESSION,side= "both")) %>%
  mutate(SEQUENCE_CLINICAL_ACCESSION=str_trim(SEQUENCE_CLINICAL_ACCESSION,side= "both"))
```

Reading GISAID File
```{r}
# wadata <- read.delim(paste0(project_folder,"/GISAID Data\\wa_gisaid_data.tsv"))
wadata <- readRDS(file=file.path(project_folder,"GISAID Data/testing_rewrite/wa_gisaid.rds"))
```

Prepping the GISAID Files
```{r}
# 
# Prep_wadata <- wadata %>% 
#   mutate(SEQUENCE_ACCESSION=str_match(Virus.name,"hCoV-19/(.*$)")[,2]) %>% 
#   select(SEQUENCE_ACCESSION, Lineage)
names(wadata)<-str_replace_all(names(wadata), c(" " = "." , "," = "" ))

# wadata2 <- wadata %>%
#   mutate(SEQUENCE_ACCESSION=str_match(Virus.name,"hCoV-19/(.*$)")[,2])
wadata2 <- wadata %>%
  mutate(SEQUENCE_ACCESSION=Virus.name)


```

Reading all files in the Missing_Lineage folder
```{r}
filenames <- list.files(paste0(project_folder,"/For_Review\\Missing_Lineage"),pattern="*.csv",full.names=TRUE)

Missing_Lineage <- data.frame()

for (i in 1:length(filenames)){
  temp_data <- read.csv(filenames[i])
  temp_data<-temp_data %>%
  mutate(SEQUENCE_STATUS=str_to_title(SEQUENCE_STATUS)) %>% 
  mutate(SEQUENCE_CLINICAL_ACCESSION=as.character(SEQUENCE_CLINICAL_ACCESSION)) %>% 
  mutate(File_PATH=filenames[i])
  Missing_Lineage <- bind_rows(Missing_Lineage, temp_data)
}
```

Deduplicating identical rows
```{r}
Missing_LineagePrep<-Missing_Lineage %>% 
  select(-File_PATH)

DedupMissing_Lineage<-distinct(Missing_LineagePrep)
```

Deduplicating against data already in WDRS
```{r}
Deduped<-anti_join(DedupMissing_Lineage, wdrs_Flat_Clean, by="SEQUENCE_ACCESSION")
```

#Connect to entire table in order to match 
```{r}
wdrs_ent <- dbGetQuery(
  connection,
  "
  SELECT Distinct CASE_ID,
    [FILLER__ORDER__NUM] as SEQUENCE_CLINICAL_ACCESSION,
    SPECIMEN__COLLECTION__DTTM as COLLECTION_DATE
  FROM [dbo].[DD_ELR_DD_ENTIRE]
  WHERE CODE = 'SARS'
  "
)
```

Match and then check for differences. There were some variable differences between ent and deduped. 
```{r}
match_ent <-
  inner_join(Deduped, wdrs_ent, by = "SEQUENCE_CLINICAL_ACCESSION")

# df_dups <- try[c("CASE_ID", "SEQUENCE_CLINICAL_ACCESSION"),]
match_ent2 <- match_ent %>%
  distinct(CASE_ID.x, SEQUENCE_CLINICAL_ACCESSION, .keep_all = TRUE)

#See if the match to entire table has any differences in CASE_ID than the original data pulled in the for_review folder. Should match
a <- match_ent2$CASE_ID.x
b <- unique(Deduped$CASE_ID)
setdiff(b, a)

c <- setdiff(b,a)
identical(c, integer(0))
# QA Check -- if there is a different between entire table and the filenames data then print a warning

if (!identical(c, integer(0))){
  print("WARNING! There is a difference in variables between entire table and data in the filenames dataset")
} else {
  print("No difference")
}

#10/27 - FA: There is a difference, need to manually add a 0 to this clinical_accession number in order for it to match to the entire table
# Deduped$SEQUENCE_CLINICAL_ACCESSION[which(Deduped$CASE_ID == )] <-
#   ""
# 
# Deduped$SEQUENCE_CLINICAL_ACCESSION[which(Deduped$CASE_ID == )]

```

After correcting differences in the data, need to check again to make sure data matches to ent table
```{r}
#Now we need to match back to the entire table and check again
matchfinal <-
  inner_join(Deduped, wdrs_ent, by = "SEQUENCE_CLINICAL_ACCESSION")

# df_dups <- try[c("CASE_ID", "SEQUENCE_CLINICAL_ACCESSION"),]
matchfinal2 <- matchfinal %>%
  distinct(CASE_ID.x, SEQUENCE_CLINICAL_ACCESSION, .keep_all = TRUE) %>%
  rename(CASE_ID = CASE_ID.x)

#See if the match to entire table has any differences in CASE_ID than the original data pulled in the for_review folder. Should match
a <- matchfinal2$CASE_ID
b <- unique(Deduped$CASE_ID)
setdiff(b, a)

c <- setdiff(b,a)
identical(c, integer(0))
# QA Check -- if there is a different between entire table and the filenames data then print a warning

if (!identical(c, integer(0))){
  print("WARNING! There is a difference in variables between entire table and data in the filenames dataset")
} else {
  print("No difference")
}

```

Found a lineage - ones that match to ent table and match to GISAID
```{r}
FoundLineage <-
  left_join(matchfinal2, wadata2, by = c("SEQUENCE_ACCESSION" = 'Virus name')) %>% filter(!is.na(Lineage)) %>%
  mutate(SEQUENCE_STATUS = if_else(Lineage == "None", "LOW QUALITY", SEQUENCE_STATUS)) %>%
  mutate(SEQUENCE_VARIANT_OPEN_TEXT = Lineage)

```

Read VOC and clean FoundLineage -- 12/2/2021 FA - I don't believe that we need to read in voc list anymore
- This hasn't been updated since the VARIANT_OPEN_TEXT field was introduced
```{r}
# voc <- readRDS(paste0(project_folder,"/Data_Objects/VOC/voc.RDS"))
Roster <- FoundLineage %>%
  # mutate(SEQUENCE_VARIANT_OPEN_TEXT = case_when(SEQUENCE_VARIANT_OPEN_TEXT %in% voc ~ SEQUENCE_VARIANT_OPEN_TEXT,
  #                                     TRUE ~ "")) %>%
  mutate(SEQUENCE_NOTES = ifelse(SEQUENCE_ACCESSION == "", "", (
    paste0(
      "Lineage identified as ",
      Lineage,
      " on ",
      today(),
      ". Lineage assignments may change over time."
    )
  ))) %>%
  select(-Lineage)
```


Pull in GISAID and CDC files to see if there are any matches between the two
```{r}
CDC_Cumulative <-
  read.csv(paste0(project_folder,"/Submissions\\CDC_Cumulative\\Washington_cumulative.csv"))

# GISAID_table <-
#   read.table(
#     file = paste0(project_folder,"/GISAID Data\\wa_gisaid_data.tsv"),
#     sep = '\t',
#     header = TRUE
#   )

# dedup records
CDC_Cumulative <- distinct(CDC_Cumulative)
wadata2 <- distinct(wadata2)
```

1. Get only CDC records that aren't in GISAID - so that we can see any that match to just CDC
2. Match to CDC
```{r}
# find records in CDC_Cum that are not in GISAID
CDC <- CDC_Cumulative %>%
  anti_join(wadata2, by = c("GISAID_name" = "Virus name"))

Prep_CDC <- CDC %>%
  mutate(SEQUENCE_ACCESSION = str_match(GISAID_name, "hCoV-19/(.*$)")[, 2]) %>%
  select(SEQUENCE_ACCESSION, lineage_PANGO_lineage)


FoundLineageCDC <-
  left_join(matchfinal2, Prep_CDC, by = ("SEQUENCE_ACCESSION")) %>%
  filter(!is.na(lineage_PANGO_lineage)) %>%
  mutate(SEQUENCE_STATUS = if_else(lineage_PANGO_lineage == "None", "LOW QUALITY", SEQUENCE_STATUS)) %>%
  mutate(SEQUENCE_VARIANT_OPEN_TEXT = lineage_PANGO_lineage)

```

Bind the rosters found from GISAID and the ones found in CDC and combine them
```{r}
if(nrow(FoundLineage)!= 0 & nrow(FoundLineageCDC)!=0){
#need to match number of columns (or add the by statement)
FoundLineage2 <- bind_rows(Roster, FoundLineageCDC) %>%
  select(
    CASE_ID,
    SEQUENCE_SGTF,
    SEQUENCE_SPECIMEN,
    SEQUENCE_REASON,
    SEQUENCE_DATE,
    SEQUENCE_LAB,
    SEQUENCE_STATUS,
    SEQUENCE_REPOSITORY,
    SEQUENCE_ACCESSION,
    SEQUENCE_VARIANT_OPEN_TEXT,
    SEQUENCE_CLINICAL_ACCESSION,
    SEQUENCE_SPECIMEN_COLLECTION_DATE,
    SEQUENCE_NOTES,
    SEQUENCE_REVIEWED,
    Case.Note
  )
} else if(nrow(FoundLineage) != 0 & nrow(FoundLineageCDC==0)){
  FoundLineage2 <- FoundLineage %>%
  select(
    CASE_ID,
    SEQUENCE_SGTF,
    SEQUENCE_SPECIMEN,
    SEQUENCE_REASON,
    SEQUENCE_DATE,
    SEQUENCE_LAB,
    SEQUENCE_STATUS,
    SEQUENCE_REPOSITORY,
    SEQUENCE_ACCESSION,
    SEQUENCE_VARIANT_OPEN_TEXT,
    SEQUENCE_CLINICAL_ACCESSION,
    SEQUENCE_SPECIMEN_COLLECTION_DATE,
    SEQUENCE_NOTES,
    SEQUENCE_REVIEWED,
    Case.Note
  )
} else if(nrow(FoundLineage)==0 & nrow(FoundLineageCDC!=0)){
  FoundLineage2 <- FoundLineageCDC %>%
  select(
    CASE_ID,
    SEQUENCE_SGTF,
    SEQUENCE_SPECIMEN,
    SEQUENCE_REASON,
    SEQUENCE_DATE,
    SEQUENCE_LAB,
    SEQUENCE_STATUS,
    SEQUENCE_REPOSITORY,
    SEQUENCE_ACCESSION,
    SEQUENCE_VARIANT_OPEN_TEXT,
    SEQUENCE_CLINICAL_ACCESSION,
    SEQUENCE_SPECIMEN_COLLECTION_DATE,
    SEQUENCE_NOTES,
    SEQUENCE_REVIEWED,
    Case.Note
  )
}
FoundLineage3 <- FoundLineage2 %>%
  mutate(
    SEQUENCE_STATUS = toupper(SEQUENCE_STATUS),
    SEQUENCE_NOTES = if_else(
      is.na(SEQUENCE_NOTES) &
        SEQUENCE_VARIANT_OPEN_TEXT != "",
      paste0(
        "Lineage identified as ",
        SEQUENCE_VARIANT_OPEN_TEXT,
        " on ",
        Sys.Date(),
        ". Lineage assignments may change over time"
      ),
      SEQUENCE_NOTES
    )
  ) %>%
  mutate(SEQUENCE_VARIANT_OPEN_TEXT = if_else(
    SEQUENCE_VARIANT_OPEN_TEXT != "" &
      !is.na(SEQUENCE_VARIANT_OPEN_TEXT),
    SEQUENCE_VARIANT_OPEN_TEXT,
    NULL
  ))%>%
  mutate(SEQUENCE_NOTES = if_else(
    SEQUENCE_STATUS=="COMPLETE",SEQUENCE_NOTES,NULL
  ))

FoundLineage4 <- FoundLineage3 %>%
  mutate(SEQUENCE_NOTES = if_else(!is.na(SEQUENCE_VARIANT_OPEN_TEXT), SEQUENCE_NOTES, NULL))

subsetmissingCaseID <- FoundLineage4 %>%
  filter(is.na(CASE_ID))

FoundLineage5 <- FoundLineage4 %>%
  filter(!is.na(CASE_ID)) %>%
  mutate(
    SEQUENCE_REASON = if_else(
      SEQUENCE_REASON == "PHL DIAGNOSTIC",
      "SENTINEL SURVEILLANCE",
      SEQUENCE_REASON
    )
  )

# SEQUENCE_NOTES = {if("SEQUENCE_NOTES" %in% names(.)) cyl else NULL})%>%
FoundLineageFinal <- FoundLineage5 %>% as_tibble() %>%
  mutate(SEQUENCE_SPECIMEN_COLLECTION_DATE = format(mdy(SEQUENCE_SPECIMEN_COLLECTION_DATE), "%m/%d/%Y")) %>%
  unique()

```






####################################################################################
## Try to match missing FAILED/LOW QUALITY lineages in flat table to 'None' Lineage in GISAID
## This could be done in the roster scripts maybe?
####################################################################################






```{r}
wdrs_flat <- dbGetQuery(
  connection,
  "
                    SELECT DISTINCT CDC_N_COV_2019_SEQUENCE_SPECIMEN, CASE_ID, CDC_N_COV_2019_SEQUENCE_ACCESSION_NUMBER,
                    CDC_N_COV_2019_SEQUENCE_CLINICAL_ACCESSION_NUMBER, CDC_N_COV_2019_SEQUENCE_REASON,
                    CDC_N_COV_2019_SEQUENCE_DATE, CDC_N_COV_2019_SEQUENCE_LAB, CDC_N_COV_2019_SEQUENCE_STATUS,
                    CDC_N_COV_2019_SEQUENCE_VARIANT_OPEN_TEXT, CDC_N_COV_2019_SEQUENCE_NOTES,
                    STATUS

                    FROM [dbo].[DD_GCD_COVID_19_FLATTENED]
                     WHERE CDC_N_COV_2019_SEQUENCE_SPECIMEN LIKE '%YES%'
                    "
)
# wdrs_Flat_Clean <-wdrs_Flat %>%
#   rename("SEQUENCE_ACCESSION"="CDC_N_COV_2019_SEQUENCE_ACCESSION_NUMBER") %>%
#   rename("SEQUENCE_CLINICAL_ACCESSION"="CDC_N_COV_2019_SEQUENCE_CLINICAL_ACCESSION_NUMBER") %>%
#   # separate_rows(SEQUENCE_ACCESSION,SEQUENCE_CLINICAL_ACCESSION, sep = ',') %>%
#   mutate(SEQUENCE_ACCESSION=str_trim(SEQUENCE_ACCESSION,side= "both")) %>%
#   mutate(SEQUENCE_CLINICAL_ACCESSION=str_trim(SEQUENCE_CLINICAL_ACCESSION,side= "both"))
#
# wdrs_Flat_Clean %>% as_tibble() %>%
#   filter(CDC_N_COV_2019_SEQUENCE_STATUS %in% c("LOW QUALITY","FAILED"))

```



```{r}
wdrs_flat_test <- wdrs_flat %>%
  separate_rows(
    CDC_N_COV_2019_SEQUENCE_SPECIMEN,
    CDC_N_COV_2019_SEQUENCE_ACCESSION_NUMBER,
    CDC_N_COV_2019_SEQUENCE_CLINICAL_ACCESSION_NUMBER,
    CDC_N_COV_2019_SEQUENCE_REASON,
    CDC_N_COV_2019_SEQUENCE_DATE,
    CDC_N_COV_2019_SEQUENCE_LAB,
    CDC_N_COV_2019_SEQUENCE_REASON,
    CDC_N_COV_2019_SEQUENCE_VARIANT_OPEN_TEXT ,
    CDC_N_COV_2019_SEQUENCE_DATE,
    CDC_N_COV_2019_SEQUENCE_STATUS,
    sep = ', ',
    convert = TRUE
  )
df <- wdrs_flat_test %>%
  mutate(
    SEQUENCE_REASON = str_trim(CDC_N_COV_2019_SEQUENCE_REASON, side = "both"),
    SEQUENCE_ACCESSION = str_trim(CDC_N_COV_2019_SEQUENCE_ACCESSION_NUMBER, side = "both"),
    SEQUENCE_CLINICAL = str_trim(CDC_N_COV_2019_SEQUENCE_CLINICAL_ACCESSION_NUMBER, side = "both"),
    SEQUENCE_DATE = str_trim(CDC_N_COV_2019_SEQUENCE_DATE, side = "both"),
    SEQUENCE_LAB = str_trim(CDC_N_COV_2019_SEQUENCE_LAB, side = "both"),
    SEQUENCE_VARIANT_OPEN_TEXT = str_trim(CDC_N_COV_2019_SEQUENCE_VARIANT_OPEN_TEXT, side = "both"),
    SEQUENCE_STATUS = str_trim(CDC_N_COV_2019_SEQUENCE_STATUS, side = "both"),
    SEQUENCE_SPECIMEN = str_trim(CDC_N_COV_2019_SEQUENCE_SPECIMEN, side = "both"),
    SEQUENCE_NOTES = str_trim(CDC_N_COV_2019_SEQUENCE_NOTES, side = "both")
  )

```


```{r}
failed <- df %>%
  filter(toupper(CDC_N_COV_2019_SEQUENCE_STATUS) %in% c("FAILED", "LOW QUALITY")) %>%
  mutate(flag = if_else(is.na(SEQUENCE_DATE) |
                          SEQUENCE_DATE == "", 1, 0))

failed %>%
  group_by(flag) %>%
  summarise(n = n())

failed <- failed %>%
  filter(flag == 0) 

failed2 <- failed%>%
  mutate(SEQUENCE_DATE = lubridate::mdy(failed$SEQUENCE_DATE))
```


```{r}
# wadata <- read.delim(paste0(project_folder,"/GISAID Data\\wa_gisaid_data.tsv"))
# metadata <-
#   read.delim(
#     paste0(project_folder,"/GISAID Data\\metadata.tsv.gz")
#   )

```




```{r}
# load	CDC Cumulative file
# CDC <-
#   read_csv(paste0(project_folder,"/Submissions/CDC_Cumulative/Washington_cumulative.csv"))
# 
# # dedup records
# CDC <- distinct(CDC)

Prep_CDC <- CDC_Cumulative %>%
  mutate(SEQUENCE_ACCESSION = str_match(GISAID_name, "hCoV-19/(.*$)")[, 2]) %>%
  filter(lineage_PANGO_lineage == "None")%>%
  mutate(
    LAB = case_when(
      str_detect(toupper(vendor), "AEGIS") ~ "Aegis",
      str_detect(toupper(vendor), "ALTIUS") ~ "Altius",
      str_detect(toupper(vendor), "ATLAS") ~ "Atlas Genomics",
      str_detect(toupper(vendor), "CENTERS FOR DISEASE") ~ "CDC",
      str_detect(toupper(vendor), "FULGENT") ~ "Fulgent Genetics",
      str_detect(toupper(vendor), "GRAVITY") ~ "Gravity Diagnostics",
      str_detect(toupper(vendor), "HELIX") ~ "Helix",
      str_detect(toupper(vendor), "INFINITY BIOLOGIX") ~ "Infinity Biologix",
      str_detect(toupper(vendor), "LABCORP|LABORATORY CORPORATION") ~ "Labcorp",
      str_detect(toupper(vendor), "OHSU|OREGON HEALTH AND SCIENCE") ~ "OHSU",
      str_detect(toupper(vendor), "WA STATE PHL|WASHINGTON STATE|WA STATE DEPARTMENT OF HEALTH") ~ "PHL",
      str_detect(toupper(vendor), "SEATTLE FLU|ATLAS|SEATTLE|NW GENOMICS|SCAN/BEDFORD|SFS/BEDFORD") ~ "NW Genomics",
      str_detect(toupper(vendor), "PROVIDENCE") ~ "Providence Swedish",
      str_detect(toupper(vendor), "QUEST") ~ "Quest",
      str_detect(toupper(vendor), "USAMRIID") ~ "USAMRIID",
      str_detect(toupper(vendor), "UW VIROLOGY|UNIVERSITY OF WASHINGTON") ~ "UW Virology",
      str_detect(toupper(vendor), "THE JACKSON LAB") ~ "The Jackson Laboratory",
      str_detect(toupper(vendor), "NAVAL HEALTH") ~ "Naval Health Research Center",
      str_detect(toupper(vendor), "FLOW DIAGNOSTICS") ~ "Flow Diagnostics",
      str_detect(toupper(vendor), "US AIRFORCE|AEROSPACE") ~ "U.S. Airforce School of Aerospace Medicine",
      str_detect(toupper(vendor), "KAISER") ~ "KP WA Research Inst",
      str_detect(toupper(vendor), "BOISE VA") ~ "Boise VA",
      str_detect(toupper(vendor), "GRUBAUGH LAB") ~ "Grubaugh Lab",
      str_detect(toupper(vendor), "LAURING LAB") ~ "Lauring Lab",
      str_detect(toupper(vendor), "ANDERSEN") ~ "Andersen",
      str_detect(toupper(vendor), "ARIZONA STATE") ~ "ASU",
      str_detect(toupper(vendor), "ATLAS") ~ "Atlas",
      str_detect(toupper(vendor), "USAMRIID") ~ "USAMRIID",
      str_detect(toupper(vendor), "FRED HUTCHINSON") ~ "Fred Hutchinson",
      str_detect(toupper(vendor), "GINKGO") ~ "Ginkgo",
      str_detect(toupper(vendor), "YALE") ~ "Yale",
      str_detect(toupper(vendor), "HYDE") ~ "Hyde",
      str_detect(toupper(vendor), "UNIVERSITY OF WASHINGTON|UW") ~ "UW Virology",
      str_detect(toupper(vendor), "OHSU") ~ "OHSU",
      TRUE ~ vendor
    ))


Prep_CDC2 <- Prep_CDC %>%
  mutate(collection_date2 = lubridate::ymd(collection_date))
```



```{r}
joinflatCDC <-
  inner_join(
    failed2,
    Prep_CDC2,
    by = c("SEQUENCE_CLINICAL" = "ID", "SEQUENCE_DATE" = "collection_date2", "SEQUENCE_LAB" = "LAB"))

```




# %>%
# mutate(SEQUENCE_STATUS=if_else(lineage_PANGO_lineage=="None", "LOW QUALITY",SEQUENCE_STATUS)) %>%
# mutate(SEQUENCE_VARIANT_OPEN_TEXT=lineage_PANGO_lineage)

```{r}
                            
md<- wadata2 %>%
  mutate(flag = if_else(is.na(date) | str_count(date) < 10 , 1, 0))

# 
# md %>%
#   group_by(flag) %>%
#   summarise(n = n())

md2 <- md %>%
  filter(flag == 0) %>%
  mutate(date = lubridate::ymd(date))

                                
# md2$pango_lineage=="None"            
# 
# md2 %>%
#   filter(pango_lineage == "None")


md3 <- md2 %>%
  filter(pango_lineage == "None")%>%
  mutate(
    LAB = case_when(
      str_detect(toupper(submitting_lab), "AEGIS") ~ "Aegis",
      str_detect(toupper(submitting_lab), "ALTIUS") ~ "Altius",
      str_detect(toupper(submitting_lab), "ATLAS") ~ "Atlas Genomics",
      str_detect(toupper(submitting_lab), "CENTERS FOR DISEASE") ~ "CDC",
      str_detect(toupper(submitting_lab), "FULGENT") ~ "Fulgent Genetics",
      str_detect(toupper(submitting_lab), "GRAVITY") ~ "Gravity Diagnostics",
      str_detect(toupper(submitting_lab), "HELIX") ~ "Helix",
      str_detect(toupper(submitting_lab), "INFINITY BIOLOGIX") ~ "Infinity Biologix",
      str_detect(toupper(submitting_lab), "LABCORP|LABORATORY CORPORATION") ~ "Labcorp",
      str_detect(toupper(submitting_lab), "OHSU|OREGON HEALTH AND SCIENCE") ~ "OHSU",
      str_detect(toupper(submitting_lab), "WA STATE PHL|WASHINGTON STATE|WA STATE DEPARTMENT OF HEALTH") ~ "PHL",
      str_detect(toupper(submitting_lab), "SEATTLE FLU|ATLAS|SEATTLE|NW GENOMICS|SCAN/BEDFORD|SFS/BEDFORD") ~ "NW Genomics",
      str_detect(toupper(submitting_lab), "PROVIDENCE") ~ "Providence Swedish",
      str_detect(toupper(submitting_lab), "QUEST") ~ "Quest",
      str_detect(toupper(submitting_lab), "USAMRIID") ~ "USAMRIID",
      str_detect(toupper(submitting_lab), "UW VIROLOGY|UNIVERSITY OF WASHINGTON") ~ "UW Virology",
      str_detect(toupper(submitting_lab), "THE JACKSON LAB") ~ "The Jackson Laboratory",
      str_detect(toupper(submitting_lab), "NAVAL HEALTH") ~ "Naval Health Research Center",
      str_detect(toupper(submitting_lab), "FLOW DIAGNOSTICS") ~ "Flow Diagnostics",
      str_detect(toupper(submitting_lab), "US AIRFORCE|AEROSPACE") ~ "U.S. Airforce School of Aerospace Medicine",
      str_detect(toupper(submitting_lab), "KAISER") ~ "KP WA Research Inst",
      str_detect(toupper(submitting_lab), "BOISE VA") ~ "Boise VA",
      str_detect(toupper(submitting_lab), "GRUBAUGH LAB") ~ "Grubaugh Lab",
      str_detect(toupper(submitting_lab), "LAURING LAB") ~ "Lauring Lab",
      str_detect(toupper(submitting_lab), "ANDERSEN") ~ "Andersen",
      str_detect(toupper(submitting_lab), "ARIZONA STATE") ~ "ASU",
      str_detect(toupper(submitting_lab), "ATLAS") ~ "Atlas",
      str_detect(toupper(submitting_lab), "USAMRIID") ~ "USAMRIID",
      str_detect(toupper(submitting_lab), "FRED HUTCHINSON") ~ "Fred Hutchinson",
      str_detect(toupper(submitting_lab), "GINKGO") ~ "Ginkgo",
      str_detect(toupper(submitting_lab), "YALE") ~ "Yale",
      str_detect(toupper(submitting_lab), "HYDE") ~ "Hyde",
      str_detect(toupper(submitting_lab), "UNIVERSITY OF WASHINGTON|UW") ~ "UW Virology",
      str_detect(toupper(submitting_lab), "OHSU") ~ "OHSU",
      TRUE ~ submitting_lab
    ))

# md3 %>%
#   group_by(LAB)%>%
#   summarise(n=n())%>%
#   view
# 
# md3 %>%
#   filter(LAB=="CDC")%>%
#   select(date,LAB)%>%
#   view
```


```{r}
joinflatmeta <-
  inner_join(
    failed2,
    md3,
    by = c("SEQUENCE_DATE" = "date", "SEQUENCE_LAB" = "submitting_lab"))

```

#One df that has all matched FAILED/LOW QUALITY to 'None' GISAID lineages
```{r}
join_fail_none <- bind_rows(joinflatCDC,joinflatmeta) 

if(nrow(FoundLineageFinal)!= 0 & nrow(join_fail_none) != 0){
FoundLineageFinal <- bind_rows(FoundLineageFinal,join_fail_none)
}else if(nrow(FoundLineageFinal)!= 0 & nrow(join_fail_none)==0){
  print("No FAILED records in WDRS could match to a lineage that is 'None' in GISAID")
}
```





####################################################################################
## write the matched data to be rostered
## Move files in the missing_lineages folder and then created LineageNotFound folder for more support
####################################################################################






Write Combined FoundLineages to Roster
```{r}
if(nrow(FoundLineageFinal)!=0){
fwrite(FoundLineageFinal,
          paste0(project_folder,"/write_roster_here\\",
            "/",
            "FoundLineages_",
            today(),
            ".csv"
          ))
  
  # send an email stating missing lineages have been matched and file is present in write_roster_here
  email_from <- ""
  email_to <- email_list1
  email_subj <- "Sequencing - Missing Lineages Matched"
  email_body <- "Missing lineages have been matched and sent to write_roster_here"

  # send it
  sendmailR::sendmail(from = email_from,
                    to = email_to,
                    subject = email_subj,
                    msg = email_body,
                    headers= list("Reply-To" = email_from),
                    control = list(smtpServer = ""))
  
} else{
  print("No Lineages matched to GISAID/CDC")
}
```

Anti join lineages to get the ones that weren't found and output to NotFound folder
- matchfinal2 is the data from the For_Review file that matched to entire table. Then anti join to combine found lineages to get what was Not Found
```{r}
LineageNotFound <-
  anti_join(matchfinal2, FoundLineageFinal, by = "SEQUENCE_ACCESSION")

```

#Function to create a new LineageNotFound folder and output files per lab by date
```{r}
 output <- function(lab, outlab) {
    
    
    if(!dir.exists(newday)){
      dir.create(newday)
    }
    
    LineageNotFound %>%
      filter(str_detect(toupper(LineageNotFound$SEQUENCE_LAB), toupper(lab)))%>%
      write_csv(paste0(
        newday,
        "/",
        outlab,
        "_",
        today(),
        ".csv", na = ""))
    
  }
```

# If there are no Lineages not found, do not output anything
```{r}
# Create a new folder if lineagenotfound is not NULL
if(nrow(LineageNotFound)!=0){
  
  newday <- file.path(paste0(project_folder,"/For_Review\\Missing_Lineage\\LineagesNotFound"), Sys.Date())
} else{
  print("LineageNotFound df is empty my dude")
}
  
# Output files to that folder above if there are files
if(nrow(LineageNotFound)!=0){
  mapply(output,
         lab=LineageNotFound$SEQUENCE_LAB,
         LineageNotFound$SEQUENCE_LAB)
  
  # send an email stating there are LineagesNotFound
  email_from <- ""
  email_to <- email_list2
  email_subj <- "Sequencing - LineagesNotFound"
  email_body <- "There are new files in the LineagesNotFound folder"

  # send it
  sendmailR::sendmail(from = email_from,
                    to = email_to,
                    subject = email_subj,
                    msg = email_body,
                    headers= list("Reply-To" = email_from),
                    control = list(smtpServer = ""))
  
  
} else {
  print("LineagesNotFound is NULL - All lineages matched")
}

```


Some of the data was missing a CASE ID so I'm going to output those too
```{r}
if(nrow(subsetmissingCaseID!=0)){
write_csv(subsetmissingCaseID,
          paste0(project_folder,"/For_Review/Missing_Lineage/",
            "/",
            "MissingCaseID_",
            today(),
            ".csv"
          ),
          na = "")
  print("There are files with missing case id's! This needs to be addressed in the roster scripts. They shouldn't output anything without a case id")
} else{
  print("No missing case ids")
}

```

Move all the files to the Reviewed folder
```{r}
file_move(dput(filenames),
         paste0(project_folder,"/For_Review\\Missing_Lineage\\Reviewed"))

```








