---
title: "S_dropout_records"
author: "Allie Warren"
date: "`r Sys.Date()`"
params:
    test_mode: TRUE
output: 
  html_document:
   toc: true
   toc_float: true
   toc_collapsed: true
   toc_depth: 2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F,
  message = F,
  warning = F,
  cache.path = "cache/",
  fig.pos = "htp",
  out.extra = "",
  dpi = 300
)

```

# Overview

This script pulls S-dropout records from the PHL RedCap and Surveillance dashboards. Only records that are marked as S-dropout and contain the necessary gene values for sequencing (NGene < 30, ORF1ab < 30, and SGENE = 0) are included in the output. Records are filtered for records that were newly uploaded since the last time that the script was run (which is stored as an rds in the Data_Objects folder) and records that could be matched to a record in WDRS are output in a roster format. This script was originally run on the same day as roster runs (M/W/F), but stopped running the script on 3/3/2022 as labs were no longer marking records as S-dropouts. The script could be adjusted to just use the gene values to identify these records, but this is recorded inconsistently (and in varying formats) by labs. This script may become relevant again at some point.

# Setup

## Load all libraries

```{r load all libraries}
library(tidyverse)
library(fuzzyjoin)
library(here)
library(lubridate)
library(readxl)
library(DBI)
library(odbc)
library(dtplyr)
library(fs)
```

## Set project folder
Set project folder as local folder for testing or as net drive folder if running for upload

```{r set project folder}

# By default set the project folder to the current R project folder
project_folder <- here()

# If running production mode (when the param test_mode is FALSE), read from and write to the net drive folder
if(!params$test_mode) {
  # read in file path to the folder on the net drive
  project_folder <- readRDS(here("Data_Objects//PHL_objects.RDS"))
}

```

## String cleanup functions

```{r string cleanup}
# String cleanup
# remove anything but alphanumeric characters; i.e. punctuation (does not remove spaces)
kill <- function(x) {str_replace_all(x, "[^[:alnum:]]", " ")}

# remove any instance of a space followed by a letter at the end of a string; usually an initial
destroy <- function(y) {str_replace_all(y, c("[[:space:]][[:alpha:]]{1}$"), " ")}

# remove white spaces
smash <- function(z) {str_replace_all(z, c(" "), "")}

# remove all non-alphanumeric characters, then remove white space around the edges, then remove initials following names, then remove any remaining spaces
annihilate <- function(w) {toupper(smash(destroy(str_trim(kill(w)))))}


```


## Read in variables
```{r}
# used to determine the number of rows used to identify data types in input files
type_n_rows <- 10000

# record of the last day that the script was run, used to separate out new records
last_day <- readRDS(file.path(project_folder, "Data_Objects/s_dropout_date.RDS"))

dropout_records <- ""
```

# Read in Data

Records are pulled from the RedCap and Surveillance dashboards from PHL. This data is downloaded using the sel_Dashboard_All script.

## Import REDCAP data

```{r import redcap data}

# get filepath for files containing REDCAP in the PHL Submissions folder
redcap_filename <- dir_ls(file.path(project_folder, "Submissions", "PHL")) %>%
  str_subset("REDCAP")

# stop script if no REDCAP file
if(length(redcap_filename) == 0){
   stop("No REDCAP file in Submissions/PHL folder")
}

# throw warning if there is more than one REDCAP file, then take the most recent file
# and continue
if (length(redcap_filename) > 1){
  print("More than one REDCAP file, selecting just the most recent file")
  
  # get time of file creation for each PHL file, and get order of oldest to most recent
  most_recent_file <- lapply(redcap_filename, function(x) file.info(x)$mtime) %>% purrr::reduce(c) %>% order()
  # use ordering of file dates to get the most recent file
  redcap_filename <- redcap_filename[most_recent_file[length(most_recent_file)]]
}


# define the list of columns of interest from the redcap viewer
redcap_columns <- c()

# read in REDCAP file
redcap <- read_xlsx(redcap_filename, guess_max = type_n_rows)
```


## Import Surveillance data

```{r import surveillance data}

# get filepath for files containing Surveillance in the PHL Submissions folder
surveillance_filename <- dir_ls(file.path(project_folder, "Submissions", "PHL")) %>%
  str_subset("Surveillance")

# stop script if no Surveillance file
if(length(surveillance_filename) == 0){
   stop("No Surveillance file in Submissions/PHL folder")
}

# throw warning if there is more than one Surveillance file, then take the most recent file
# and continue
if (length(surveillance_filename) > 1){
  print("More than one Surveillance file, selecting just the most recent file")
  
  # get time of file creation for each PHL file, and get order of oldest to most recent
  most_recent_file <- lapply(surveillance_filename, function(x) file.info(x)$mtime) %>% purrr::reduce(c) %>% order()
  # use ordering of file dates to get the most recent file
  surveillance_filename <- surveillance_filename[most_recent_file[length(most_recent_file)]]
}

# read in Surveillance file
surveillance <- read_xlsx(surveillance_filename, guess_max = type_n_rows)
```

# WDRS

## Connect
```{r wdrs connection}

# read in connection credentials
conn_list <- readRDS(file.path(project_folder, "Data_Objects/conn_list.RDS"))



connection <- DBI::dbConnect(odbc::odbc(), 
                             Driver = conn_list[[1]], 
                             Server = conn_list[[2]], 
                             Database = conn_list[[3]], 
                             Trusted_connection = conn_list[[4]], 
                             ApplicationIntent = conn_list[[5]])
```

## Query WDRS ELR Entire table

```{r}
wdrs_entire <- dbGetQuery(
    connection,
    "SELECT Distinct CASE_ID,
                       [PATIENT__FIRSTNAME] as FNAME,
                       [PATIENT__LASTNAME] as LNAME,
                       [FILLER__ORDER__NUM] as SpecimenId,
                       [PATIENT__DATE__OF__BIRTH] as DOB_WDRS,
                       [SPECIMEN__COLLECTION__DTTM] as COLLECTION_DATE_WDRS,
                       [FIRST_NAME] as FIRST_NAME,
                       [LAST_NAME] as LAST_NAME
  FROM [dbo].[DD_ELR_DD_ENTIRE] WHERE CODE = 'SARS'
  AND STATUS != '6'
  AND WDRS__RESULT__SUMMARY != 'NEGATIVE'"
  ) 
  

  
  # clean up WDRS names, remove characters that are not alphanumeric, remove white spaces
  wdrs_entire <- wdrs_entire %>%
    mutate(FNAME = annihilate(wdrs_entire$FNAME),
           LNAME = annihilate(wdrs_entire$LNAME),
           SpecimenId = as.character(SpecimenId),
           FIRST_NAME = annihilate(wdrs_entire$FIRST_NAME),
           LAST_NAME = annihilate(wdrs_entire$LAST_NAME)) %>%
    unite(NAME_WDRS,
          c(FNAME, LNAME),
          remove = TRUE,
          na.rm = TRUE) %>%
    unite(ALT_NAME_WDRS,
          c(FIRST_NAME, LAST_NAME),
          remove = TRUE,
          na.rm = TRUE)


```

## Create separate records for alternate names
Some WDRS records have two possible names (different spellings, nicknames etc) within record, and, for the RedCap data we match on 
name and DOB (since the Ids are unreliable), so we want to allow matches with either version of the name

```{r wdrs names}
# Combine PATIENT NAME and NAME columns
wdrs_entire <- wdrs_entire %>%
   mutate(NAME_WDRS = case_when(
      NAME_WDRS == "" ~ ALT_NAME_WDRS,
      NAME_WDRS != "" ~ NAME_WDRS)
      )

# Get records where the PATIENT NAME and NAME columns differ
alt_wdrs_name <- filter(wdrs_entire, NAME_WDRS != ALT_NAME_WDRS) %>%
  mutate(NAME_WDRS = ALT_NAME_WDRS)

# Add rows using the alternate name, duplicating the other columns
wdrs_entire <- rbind.data.frame(wdrs_entire, alt_wdrs_name)


```

# S-Dropout records

## Create output data frame
```{r create empty df}

S_dropout_matched <- as.data.frame(matrix(data=0, nrow=0, ncol=2)) %>% magrittr::set_colnames(c("SEQUENCE_SGTF",
                                                                                                "CASE_ID"))
```

## Filter RedCap for relevant records
```{r filter and select redcap}
# Records no longer being marked as S-Dropout, would need to change this logic to 
# catch these records (and if changing this logic without changing the last_day
# it won't catch older records)
redcap_s_dropout <- filter(redcap, toupper(ProjectOption) == 'S-DROPOUT') %>%
  # identify column containing relevant genes and populate new column with corresponding value
  mutate(NGENE = case_when((str_detect(annihilate(Target1), "NGENE") | annihilate(Target1) == "N") ~ Ct1,
                           (str_detect(annihilate(Target2), "NGENE") | annihilate(Target2) == "N") ~ Ct2,
                           (str_detect(annihilate(Target3), "NGENE") | annihilate(Target3) == "N") ~ Ct3,
                           (str_detect(annihilate(Target4), "NGENE") | annihilate(Target4) == "N") ~ Ct4),
         ORF1ab = case_when(str_detect(annihilate(Target1), "ORF") ~ Ct1,
                           str_detect(annihilate(Target2), "ORF") ~ Ct2,
                           str_detect(annihilate(Target3), "ORF") ~ Ct3,
                           str_detect(annihilate(Target4), "ORF") ~ Ct4),
         SGENE = case_when((str_detect(annihilate(Target1), "SGENE") | annihilate(Target1) == "S") ~ Ct1,
                           (str_detect(annihilate(Target2), "SGENE") | annihilate(Target1) == "S") ~ Ct2,
                           (str_detect(annihilate(Target3), "SGENE") | annihilate(Target1) == "S") ~ Ct3,
                           (str_detect(annihilate(Target4), "SGENE") | annihilate(Target1) == "S") ~ Ct4)) %>%
  # populate column if gene names are not specified
  mutate(NGENE = case_when((is.na(NGENE) & is.na(Target1)) ~ Ct1,
                           TRUE ~ NGENE),
         ORF1ab = case_when((is.na(ORF1ab) & is.na(Target2)) ~ Ct2,
                           TRUE ~ ORF1ab)) %>%
  # convert gene values to numeric
  mutate(NGENE = as.numeric(NGENE),
         ORF1ab = as.numeric(ORF1ab),
         SGENE = as.numeric(SGENE)) %>%
  # filter to records that fit gene value criteria for sequencing
  filter(NGENE < 30 | is.na(NGENE),
         ORF1ab < 30 | is.na(ORF1ab),
         SGENE == 0 | is.na(SGENE) | SGENE == "") %>%
  # clean up name and DOB
  mutate(NAME_REDCAP = paste0(annihilate(FirstName), "_", annihilate(LastName)),
         DOB_REDCAP = as.Date(BirthDate)) %>%
  select("FirstName",
         "LastName",
         "BirthDate",
         "CollectionDate",
         "UploadedDate",
         "ProjectOption",
         "NAME_REDCAP",
         "DOB_REDCAP",
         "AccessionNumber",
         "SubmissionDate",
         "NGENE",
         "ORF1ab") %>%
  mutate(Upload_Date =  ymd_hms(UploadedDate))

redcap_s_dropout_recent <- filter(redcap_s_dropout, Upload_Date > last_day) %>% 
  select(-Upload_Date)


```

## Match RedCap to WDRS
```{r match redcap to WDRS and write to file}

if(nrow(redcap_s_dropout_recent) > 0) {
  # Join on name and DOB - only allowing for exact matches
  # A Fuzzy matching process could also be added here
  redcap_wdrs <- inner_join(redcap_s_dropout_recent, wdrs_entire, by = c("NAME_REDCAP" = "NAME_WDRS", "DOB_REDCAP" = "DOB_WDRS"))
  redcap_wdrs <- redcap_wdrs[!duplicated(redcap_wdrs[,c('CASE_ID', 'AccessionNumber')]),]
  
  # Separate out perfect matches
    # collection dates must be within 14 days
  redcap_matches <- redcap_wdrs %>%
    mutate(int = interval(
      ymd(CollectionDate),
      ymd(COLLECTION_DATE_WDRS)
    ),
    COLLECTION_DATE_DISTANCE = abs(time_length(int, unit = "day")))
  
  redcap_matches <- filter(redcap_matches, COLLECTION_DATE_DISTANCE <= 14,
                           NGENE < 30,
                           ORF1ab < 30)
  
    # matches to only one WDRS record
  redcap_matches <- redcap_matches[!(duplicated(redcap_matches$AccessionNumber) | duplicated(redcap_matches$AccessionNumber, fromLast = T)),]
  
  # get redcap not perfect matches, which are added to a review file for manual matching
  redcap_review <- filter(redcap_s_dropout_recent, !AccessionNumber %in% redcap_matches$AccessionNumber) %>%
    select("FirstName",
         "LastName",
         "BirthDate",
         "CollectionDate",
         "UploadedDate",
         "ProjectOption",
         'SubmissionDate')
  
  # create output dataframe
  if(nrow(redcap_matches) > 0) {
    S_dropout_matched <- rbind(S_dropout_matched, redcap_matches %>%
      mutate(SEQUENCE_SGTF = "YES") %>%
      select(CASE_ID, 
             SEQUENCE_SGTF))
  }
  
  if(nrow(redcap_review) > 0) {
    r_review_filepath <- file.path(project_folder, "S-Dropouts", paste0("redcap_review_s_dropouts_", today(), ".csv"))
  write_csv(redcap_review, r_review_filepath)
  
  dropout_records <- paste0(dropout_records, "\n A file containing ", nrow(redcap_review), " S-Dropout records from RedCap that require review/manual matching has been written to the folder. The file can be found at: ", "\n", r_review_filepath, "\n")
  }


}

```

##  Filter Surveillance for relevant records
```{r filter and select surveillance}
# Records are no longer being marked as S-Dropout. With the current inputs we would 
# need to change this logic to identify S-dropout records (and if changing this logic without changing the last_day
# it won't catch older records)
surveillance_s_dropout <- filter(surveillance, toupper(Reason) == 'S-DROPOUT',
                                 SGene == 0 | SGene == "" | is.na(SGene) | SGene == "-",
                                 NGene < 30,
                                 ORF1ab < 30) %>%
  select("AccessionId",
         "CollectionDate",
         "UploadedDate",
         "Reason",
         "SubmissionDate") %>%
  mutate(Upload_Date = ymd_hms(UploadedDate))

# Inlcude only records that are new since the last time the script was run
surveillance_s_dropout_recent <- filter(surveillance_s_dropout, Upload_Date > last_day) %>% 
  select(-Upload_Date)


```

## Match Surveillance to WDRS

```{r match surveillance to WDRS and write to file}
if(nrow(surveillance_s_dropout_recent) > 0) {

    # match to WDRS using Accession Id
  surveillance_wdrs <- inner_join(surveillance_s_dropout_recent, wdrs_entire, by = c('AccessionId' = 'SpecimenId'))
  
  surveillance_wdrs <- surveillance_wdrs[!duplicated(surveillance_wdrs[,c('CASE_ID', 'AccessionId')]),]
    
  # Separate out perfect matches
    # collection dates must be within 14 days
  surveillance_matches <- surveillance_wdrs %>%
    mutate(int = interval(
      ymd(CollectionDate),
      ymd(COLLECTION_DATE_WDRS)
    ),
    COLLECTION_DATE_DISTANCE = abs(time_length(int, unit = "day")))
  
  surveillance_matches <- filter(surveillance_matches, COLLECTION_DATE_DISTANCE <= 14)
  
    # matches to only one WDRS record
  surveillance_matches <- surveillance_matches[!(duplicated(surveillance_matches$AccessionId) | duplicated(surveillance_matches$AccessionId, fromLast = T)),]
  
  # get surveillance not perfect matches, which are written to a review file for manual matching
  surveillance_review <- filter(surveillance_s_dropout_recent, !AccessionId %in% surveillance_matches$AccessionId) %>%
    select("AccessionId",
         "CollectionDate",
         "UploadedDate",
         "Reason",
         "SubmissionDate")
  
  # write to files
  if(nrow(surveillance_matches) > 0) {
    S_dropout_matched <- rbind(S_dropout_matched, surveillance_matches %>%
      mutate(SEQUENCE_SGTF = "YES") %>%
      select(CASE_ID,
             SEQUENCE_SGTF))
    }
  
  if(nrow(surveillance_review) > 0) {
    s_review_filepath <-  file.path(project_folder, "S-Dropouts", paste0("surveillance_review_s_dropouts_", today(), ".csv"))
  write_csv(surveillance_review, s_review_filepath)
  
  dropout_records <- paste0(dropout_records, "\n A file containing ", nrow(surveillance_review), " S-Dropout records from Sentinel Surveillance that require review/manual matching has been written to the folder. The file can be cound at: ", "\n", s_review_filepath, "\n")
  }

}

```

# Write output

Write output file, in roster format for records that could be matched
```{r write to file}
if(nrow(S_dropout_matched) > 0) {
  # Compile into roster format
  s_dropout_roster <- 
      S_dropout_matched %>%
      mutate(SEQUENCE_SGTF = 'Yes',
             SEQUENCE_SPECIMEN = "",
             SEQUENCE_REASON = "",
             SEQUENCE_DATE = "",
             SEQUENCE_LAB = "",
             SEQUENCE_STATUS = "",
             SEQUENCE_REPOSITORY = "",
             SEQUENCE_ACCESSION = "",
             SEQUENCE_VARIANT_OPEN_TEXT = "",
             SEQUENCE_CLINICAL_ACCESSION = "",
             SEQUENCE_SPECIMEN_COLLECTION_DATE = "",
             SEQUENCE_NOTES = "",
             SEQUENCE_REVIEWED = "",
             Case.Note = 'External data question package updated by COVID19 Sequencing Roster.') %>%
      select(CASE_ID,
             SEQUENCE_SGTF,
             SEQUENCE_SPECIMEN,
             SEQUENCE_REASON,
             SEQUENCE_DATE,
             SEQUENCE_LAB,
             SEQUENCE_STATUS,
             SEQUENCE_REPOSITORY,
             SEQUENCE_ACCESSION,
             SEQUENCE_VARIANT_OPEN_TEXT,
             SEQUENCE_CLINICAL_ACCESSION,
             SEQUENCE_SPECIMEN_COLLECTION_DATE,
             SEQUENCE_NOTES,
             SEQUENCE_REVIEWED,
             Case.Note)

  # remove duplicate rows
  s_dropout_roster <- s_dropout_roster[!duplicated(s_dropout_roster$CASE_ID),]
  

  matched_file_path <- file.path(project_folder, "S-Dropouts", paste0("matched_s_dropouts_", today(), ".csv"))
  write_csv(s_dropout_roster, matched_file_path)
  
  dropout_records <- paste0(dropout_records, "\n A file containing ", nrow(s_dropout_roster), " S-Dropout records matched exactly to WDRS records (using name and DOB or by matching AccessionId to SpecimenId) has been written to the folder. The file can be found at: ", "\n", matched_file_path, "\n")
}

```

# Update date

Record date/time that script completed 
```{r update date}

write_rds(ymd_hms(now(), tz='UTC'), file.path(project_folder, "Data_Objects/s_dropout_date.RDS"))

```

# Delete files
If running this script more frequently, delete the input files
If only running the script on M/W/F then these same files will be used by the Dashboard_All script (which will then move the files to the Completed folder)
```{r delete files}
# if not a day to roster data, delete the files
# if((!wday(today()) %in% c(2,4,6)) & !params$test_mode) {
#   # get PHL file path
#   phl_filename <- dir_ls(file.path(project_folder, "Submissions", "PHL")) %>%
#     str_subset(paste0("PHL/PHL_", today()))
#   
#   # delete all files
#   file_delete(redcap_filename)
#   
#   file_delete(surveillance_filename)
#   
#   file_delete(phl_filename)
# }
```

# Email
```{r send email}
total_records <- nrow(redcap_s_dropout_recent) + nrow(surveillance_s_dropout_recent)

# assign email components to vectors
email_from <- ""
email_to <- c("", 
              "")
email_subject <- paste0("S-Dropout Records ", month(today()), "/", day(today()))
email_body <- paste0("The PHL Records have been filtered for S-Dropout records for ", today(), ". ", total_records, " S-Dropout records were identified. ", "\n\n", dropout_records)

# send it only if running in production mode, where data is read from and written to the net drive
if(!params$test_mode) {
  sendmailR::sendmail(from = email_from,
                      to = email_to,
                      subject = email_subject,
                      msg = email_body,
                      headers= list("Reply-To" = email_from),
                      control = list(smtpServer = "")
                      )
}
```



